---
title: "N-rate Timing"
date: "`r Sys.Date()`"
# format: pdf
# format: revealjs
# format:
  # docx:
    # reference-doc: word_style.docx
editor: visual
execute: 
  echo: false
  message: false
  warning: false
editor_options: 
  chunk_output_type: console
---

### To do

Distill brainstorming done in the EDA scripts, report.Rmd and Dominics word doc into a singular first draft.

# Data

```{r setup}
library(dplyr)
library(tidyverse)
library(multcomp)
library(emmeans)

read.csv(
  "data_nrate-all.csv"
) %>% 
  as_tibble() %>% 
  mutate(stand.age = ordered(stand.age)) %>% 
  mutate(
    across(
      c(block,location, id),
      factor
    )
  ) %>% 
  mutate(timing = gsub("[0-9]","",treatment)) %>% 
  mutate(n.total = fall+spring+summer) -> dat

```

```{r print datasets, include=F}
dat %>% 
  slice(1:6) %>% 
  knitr::kable(
    caption = "first 5 lines of dataset"
  )

read.csv(
  "metadata_nrate-all.csv"
) %>% 
  knitr::kable(
    capation= "metadata"
  )
```

# Methods

```{r, include=F}
read.csv(
  "field_activities.csv"
) %>% 
  knitr::kable(
    caption = "Dates of key field activities"
  )
```

# Lodging

Understanding lodging is not a primary objective of this experiment, this data was only collected as a covariate if there happened to be a lot of lodging.

```{r}
dat %>% 
  ggplot(aes(lodging)) +
  stat_bin() + 
  facet_wrap(~location)
```

Only R100 and V17 showed lodging, and only R100 had severe lodging to the point where the yield data probably is not very accurate. \## R100

R100 only had N applied in the second and third year.

```{r}
dat %>% 
  filter(location == "R100") %>% 
  ggplot(aes(stand.age,lodging)) +
  geom_boxplot()
```

Lodging only occurred in the stands third year of production.

```{r}
dat %>% 
  filter(location == "R100") %>% 
  ggplot(aes(fall+spring+summer,lodging)) +
  geom_point() +
  stat_smooth(
    method = "lm",
    formula = y~log(x+1),
    se=F
  )
```

We observe a general increase in lodging as nitrogen rate increases. We fit a logistic curve since lodging cannot be greater than 10 and we expect as we increase N rate more lodging will get closer to 10. This curve is obviously not perfect.

```{r}
dat %>% 
  filter(location == "R100" & stand.age =="3") %>% 
  ggplot(aes(fall+spring+summer,lodging, col = timing)) +
  geom_point()
```

Looking at treatment timing with N rate, I see a bit more red dots (fall) with lower lodging, but it's messy. Sprint split has some of the most lodging, but also an outlier at near 150 kg N ha where there was no lodging

```{r}
dat %>% 
  filter(location == "R100" & stand.age =="3") %>% 
  ggplot(aes(fall+spring+summer,lodging, col = timing)) +
  stat_summary() +
  stat_smooth(
    se=F,
    method = "lm",
    formula = y~log(x+1)+0
  ) +
  labs(
    caption = "fall split line is hidden underneath spring split"
  )
```

Here we are fitting a logistic regression with a y intercept of zero because we assume at 0N there is no lodging (as shown with control plots) and that lodging score will increase as nitrogen rate increases but that lodging will never exceed 10. The takeaway from this figure is that there is no lodging at 0N and that you see less lodging when you apply in fall and more when you apply in spring and summer.

```{r}
dat %>% 
  filter(location == "R100" &
           stand.age == "3") %>% 
  pivot_longer(
    cols = c(fall,spring,summer),
    values_to = "n.rate",
    names_to = "n.timing"
  ) %>% 
  # glimpse() #lodging is now repeated 3 times
  ggplot(aes(x=n.rate,lodging,
             group = n.timing,
             col = n.timing)) +
  stat_summary() +
  stat_smooth(
    se=F,
    method = "lm"
  )
```

This is a funkier way to look at this data based on how I organized the data and it is confusing so may not be worth trying to interpret.. Every lodging observation has a corresponding amount of N applied in fall, spring and summer. All of the spring plots, for example, also are plots where no N was applied in the fall. So within each of the timings (fall, spring or summer), the entire lodging dataset is represented. This also means the lodging dataset is duplicated 3 times when we do this. I am not sure if this is ok, but this is how I think this can be interpreted.

With more N there is more lodging, but we also see more lodging when the N is applied closer to harvest.

| Timing of N | m (lodging \~ m\*nrate + b) |
|-------------|-----------------------------|
| fall        | -0.002                      |
| spring      | 0.025                       |
| summer      | 0.041                       |

: Linear regression parameters for lodging response to nitrogen timing

```{r, eval=F}
dat %>% 
  filter(location == "R100" &
           stand.age == "3") %>% 
  pivot_longer(
    cols = c(fall,spring,summer),
    values_to = "n.rate",
    names_to = "n.timing"
  ) %>% 
  filter(
    n.timing=="fall"
    # n.timing == "spring"
    # n.timing == "summer"
  ) %>% 
  lm(
    lodging~n.rate,.
  )
```

### Analysis

When we fit logistic curves, it looked like the lodging was lowest in the control and fall plots. Let's see if this is true. 

```{r}
dat %>% 
  filter(location == "R100" & stand.age =="3") %>% 
  lm(
    formula = lodging~log(n.total+1)*timing+0,.
  )  %>% 
  anova()
```

We cannot reject Ho that timing has an impact on lodging.

```{r}

dat %>% 
  filter(location == "R100" & stand.age =="3") %>% 
  lm(
    formula = lodging~timing,.
  )  %>% 
  # anova() %>% 
  emmeans::emmeans(~timing) %>% 
  cld()
```

Obviously, if we look just at timing then the control plots have less lodging than the rest of the plots receiving N. 
Takeaway here is that lodging is correlated with increases in N, but we cannot reject the Ho that timing does not have an impact. It looks like fall applied N had less impact on lodging, but we cannot reject Ho. 

Based on analysis, we are better off just fitting a general curve to lodging by nrate, we will need to do more analysis to find the best curve with the most fitting relationship. 


