---
title: "N-rate Timing"
subtitle: "first draft"
date: "`r Sys.Date()`"
# format: pdf
# format: revealjs
# format:
#   docx:
#     reference-doc: word_style.docx
editor: visual
execute: 
  echo: false
  message: false
  warning: false
editor_options: 
  chunk_output_type: console
---

# Key points

- We applied N at different rates and timings to IWG stands over 10 site-years

- We observed serious lodging in only 1 site-year, and observed lower lodging when N was applied at lower rates and further away from harvest (i.e. fall vs. spring or summer). Lodging was negatively correlated with plant height ($r=-0.5$)

- For both cumulative and yearly yield response to nitrogen rate, a quadratic relationship fit the data best, suggesting there is an optimal N rate.

- Combined across all site-years, yields were higher when N was applied in the fall versus in the spring or when split with a summer application. 

- Of sites where we tracked yield over three years, we failed to reject the Ho that yields differed among stands that recieved different N rates and timings. 

- The relationship between N rate and timing is complex to measure in field experiments. Site conditions can greatly change the amount of N that becomes available to the plant, especially when applied as urea on the soil surface. More research is required to capture the variability of yield response to N rate and timing. 

- The optimal fertility program is likely site specific. 

# Data

```{r setup}
library(dplyr)
library(tidyverse)
library(multcomp)
library(emmeans)
library(lme4)
library(car)
theme_set(theme_bw())

read.csv(
  "data_nrate-all.csv"
) %>% 
  as_tibble() %>% 
  mutate(stand.age = ordered(stand.age)) %>% 
  mutate(
    across(
      c(block,location, id),
      factor
    )
  ) -> dat

```

```{r print datasets, include=F}
dat %>% 
  slice(1:6) %>% 
  knitr::kable(
    caption = "first 5 lines of dataset"
  )

read.csv(
  "metadata_nrate-all.csv"
) %>% 
  knitr::kable(
    capation= "metadata"
  )
```

# Methods

```{r field activities, include=F}
read.csv(
  "field_activities.csv"
) %>% 
  knitr::kable(
    caption = "Dates of key field activities"
  )
```

```{r total dataset, include=F}
dat %>% 
  group_by(n.total,timing,stand.age,location) %>% 
  tally() %>% 
  # print(n=120)
  knitr::kable(
    caption = "The n-rate timing dataset"
  )
```

Our dataset is unbalanced because design was not consistent across sites. First we try to see a consistent response across site-years, then we may analyze each site seperately. 

```{r}
dat %>% 
  group_by(location,year) %>% 
  tally() %>% 
  knitr::kable(
    caption = "We have 10 site-years, more than enough to be a random effect"
  )
```


# Lodging

Understanding lodging is not a primary objective of this experiment, this data was only collected as a covariate if there happened to be a lot of lodging.

In general, if lodging is above 6, the yield data is questionable.

```{r, include=F}
dat %>% 
  ggplot(aes(lodging)) +
  stat_bin() + 
  facet_wrap(~location)
```

Only R100 and V17 showed lodging, and only R100 had severe lodging to the point where the yield data probably is not very accurate.

## R100

R100 only had N applied in the second and third year.

```{r, include=F}
dat %>% 
  filter(location == "R100") %>% 
  ggplot(aes(stand.age,lodging)) +
  geom_boxplot()
```

Lodging only occurred in the stands third year of production.

```{r}
dat %>% 
  filter(location == "R100") %>% 
  ggplot(aes(fall+spring+summer,lodging)) +
  geom_point() +
  stat_smooth(
    method = "lm",
    formula = y~log(x+1),
    se=F
  ) +
  labs(
    x="N rate (kg N ha)",
    y="Lodging score (0=no lodging)"
  )
```

We observe a general increase in lodging as nitrogen rate increases. We fit a logistic curve since lodging cannot be greater than 10 and we expect as we increase N rate more lodging will get closer to 10. This curve is obviously not perfect.

```{r, include=F}
dat %>% 
  filter(location == "R100" & stand.age =="3") %>% 
  ggplot(aes(fall+spring+summer,lodging, col = timing)) +
  geom_point()
```

<!-- Looking at treatment timing with N rate, I see a bit more red dots (fall) with lower lodging, but it's messy. Sprint split has some of the most lodging, but also an outlier at near 150 kg N ha where there was no lodging -->

```{r}
dat %>% 
  filter(location == "R100" & stand.age =="3") %>% 
  ggplot(aes(fall+spring+summer,lodging, col = timing)) +
  stat_summary() +
  stat_smooth(
    se=F,
    method = "lm",
    formula = y~log(x+1)+0
  ) +
  labs(
    caption = "fall split line is hidden underneath spring split",
    x="N rate (kg N ha)",
    y="loding score (0=no lodging)"
  )
```

Here we are fitting a logistic regression with a y intercept of zero because we assume at 0N there is no lodging (as shown with control plots) and that lodging score will increase as nitrogen rate increases but that lodging will never exceed 10. The takeaway from this figure is that there is no lodging at 0N and that you see less lodging when you apply in fall and more when you apply in spring and summer.

### Analysis

```{r}
dat %>% 
  filter(location == "R100" & stand.age =="3") %>% 
  lm(
    formula = lodging~log(n.total+1)*timing+0,.
  )  %>% 
  anova()
```

We reject Ho that the rate of nitrogen does not impact lodging. 

We fail to reject Ho that timing has an impact on lodging.

# Plant height

Plant height is also not a measurement of primary interest.

To what extent does plant height relate to lodging in R100?

## R100 year 3

```{r}
dat %>% 
  filter(location == "R100" & stand.age =="3") %>% 
  ggplot(aes(fall+spring+summer,height)) +
  geom_point() +
  geom_smooth(
    se=F,
    method = "lm",
    formula = y~poly(x,2)
  )
```

We observe an overall trend of decreasing plant height as N rate increases, modeled best quadratically.

```{r}
dat %>% 
  filter(location == "R100" & stand.age =="3") %>% 
  lm(
    height~poly(n.total,2)+timing,.
  ) %>% 
  anova()
```

We reject Ho that nrate does not impact height

We fail to reject Ho that timing has no effect on plant height

We have learned from R100 in it's third stand age that as nrate increases, there is an increase in lodging and a decrease in plant height. To what extent are they correlated?

```{r}
dat %>% 
  filter(location == "R100" & stand.age =="3") %>%
  dplyr::select(lodging,height) %>% 
  cor(method = "pearson") 
```

We observe a pearson correlation coefficient of -0.5 between height and lodging. This is considered between a moderate and strong correlation.

# Yield

## cumulative

Cumulative yield of kernza stands after 3 years of N fertilizer. We are subsetting dataset, Only V17 and Staples meet this criteria (6 site years). We sum across stand.age to create a cumulative yield and a cumulative amount of N applied, then divide both values by 3 to get a yearly yield~N response. 

```{r, include=F}
dat %>% 
  # Removing R100 plots in third year because lodging was so high
  filter(location != "R100" | stand.age != "3") %>% 
  filter(stand.age=="3") %>% 
  ggplot(aes(yield.cum)) +
  stat_bin()
```

### quadratic linear model, yield response to N

```{r}

dat %>% 
  filter(location != "R100" | stand.age != "3") %>% 
  filter(stand.age=="3") %>% 
  lm(
    yield.cum ~ poly(cumn,1),.
  ) %>% 
  AIC()

dat %>% 
  filter(location != "R100" | stand.age != "3") %>% 
  filter(stand.age=="3") %>% 
  lm(
    yield.cum ~ poly(cumn,2),.
  ) %>% 
  AIC()

# quadratic is better than linear for fitting the data

dat %>% 
  filter(location != "R100" | stand.age != "3") %>% 
  filter(stand.age=="3") %>% 
  lm(
    yield.cum~poly(cumn,2),.
  ) %>% 
  anova()

dat %>% 
  filter(location != "R100" | stand.age != "3") %>% 
  filter(stand.age=="3") %>% 
  ggplot(aes(cumn/3,yield.cum/3)) +
  geom_point() +
  geom_smooth(
    se=F,
    method = "lm",
    formula = y~poly(x,2)
  ) +
  labs(
    x="N rate (kg N ha)",
    y="Yield (kg ha)"
  )
```

We reject the Ho that N rate does not impact yield

Quadratic model provides the best fit

```{r}
dat %>% 
  filter(location != "R100" | stand.age != "3") %>% 
  filter(stand.age=="3") %>% 
  filter(timing == "Control" |
           timing == "Fall" |
           timing == "Spring") %>% 
  lm(
    yield.cum ~ poly(cumn,2)*location*timing,.
  ) %>% 
  anova()
```

We fail to reject the Ho that timing does not impact yield

### quadratic linear mixed effect model

Here we have our fixed effect of cumulative N, timing and a random effect of block. Since we only have two sites, location is treated as a fixed effect. 

```{r}
dat %>% 
  filter(location != "R100" | stand.age != "3") %>% 
  filter(stand.age=="3") %>% 
  filter(timing == "Control" |
           timing == "Fall" |
           timing == "Spring") %>% 
  lmer(
    yield.cum ~ poly(cumn,2)*timing*location+(1|block),.
  ) %>% 
  # summary()
  # AIC()
  car::Anova()
```

We fail to reject Ho that yield does not differ N rate or timing

```{r}
dat %>% 
  filter(location != "R100" | stand.age != "3") %>% 
  filter(stand.age=="3") %>% 
  filter(timing == "Control" |
           timing == "Fall" |
           timing == "Spring") %>% 
  ggplot(aes(cumn/3,yield.cum/3,
             col=location)) +
  geom_point() +
  geom_smooth(
    # aes(col=location),
    se=F,
    method="lm",
    formula = y~y) +
  labs(
    x="N rate (kg N ha)",
    y="Yield (kg ha)"
  )
```

TAKEAWAY: We applied N at differing rates and timings over 3 years at two locations. We cannot reject the Ho that the amount of N and the timing of N do not impact the cumulative yield over the 3 years when modelled as a fixed effect model. Personally, I would say our data suggests at around 60 kg N ha per year results in best grain yields and then adding more N has no effect. When modeled as a simple quadratic linear model, we can make this conclusion, but when modelled as a mixed effect model we cannot. 

## Yearly performance

We are using all site years except third stand age of R100 due to high lodging. 

How does N timing and N amount correlate with yield in a given year?

```{r, include=F}
dat %>% 
  filter(location != "R100" | stand.age != "3") %>% 
  ggplot(aes(yield)) +
  stat_bin()
```

### full model, site-years as random

```{r, include=F}
dat %>% 
  filter(location != "R100" | stand.age != "3") %>%
  group_by(location,year) %>%
  tally()

# We still have 9 site-years, let's model these as random and see if we see an effect of nrate

dat %>% 
# removing yield data from lodged plots
  filter(location != "R100" | stand.age != "3") %>%
# removing instances where there was a Fall timing but no N was applied
  filter(timing !="Fall" | fall != 0) %>% 
  lmer(
    yield~poly(n.total,2)*timing*stand.age + (1|location/year/block),.
  ) %>% 
  # summary()
  car::Anova()

# we know from this model that there are no interactions between fixed effects and that it is rank defficient because we are trying to predict so many interactions on so little data. We rerun the model without the interaction terms

```

We reject Ho that stand.age, timing and nitrogen rate do not impact yield 

Here we have subsetted data so we have removed instances where a timing was Fall but no fall N was applied. We only start doing this here because this is the first time we are looking at timing across years at each year. 

#### nitrogen rate on yield

```{r nrate effect}
dat %>% 
# removing yield data from lodged plots
  filter(location != "R100" | stand.age != "3") %>%
# removing instances where there was a Fall timing but no N was applied
  filter(timing !="Fall" | fall != 0) %>% 
  filter(timing!="Fall-Spring Split") %>% 
  ggplot(aes(n.total,yield)) +
  geom_point(
    # aes(col=timing)
  ) +
  geom_smooth(
    method = "lm",
    se=F,
    formula = y~poly(x,2)
  ) +
  labs(
    x="N rate (kg N ha)",
    y="Yield (kg ha)"
  )
```

We previously learned the relationship between N rate and yield is best modelled quadratically and then we rejected Ho that nitrogen rate does not impact yield. Here we are visualizing the subsetted data used in the model. 

#### timing on yield

```{r timing effect}
dat %>% 
# removing yield data from lodged plots
  filter(location != "R100" | stand.age != "3") %>%
# removing instances where there was a Fall timing but no N was applied
  filter(timing !="Fall" | fall != 0) %>% 
  lmer(
# changing random effect structure due to singular fit
    yield~poly(n.total,2)+timing+stand.age + (1|location/year/block),.
  ) %>% 
  # summary()
  # car::Anova()
  emmeans(~timing) %>% 
  cld(
    Letters = letters,
    reverse = T
  ) %>% 
  mutate(group = str_trim(.group)) ->tuk1

dat %>% 
# removing yield data from lodged plots
  filter(location != "R100" | stand.age != "3") %>%
# removing instances where there was a Fall timing but no N was applied
  filter(timing !="Fall" | fall != 0) %>% 
  group_by(timing) %>% 
  tally() %>% 
  left_join(tuk1) %>% 
  mutate(
    CI = paste0(round(lower.CL),"-",round(upper.CL))
  ) %>% 
  arrange(desc(emmean)) %>% 
  mutate(emmean = paste0(round(emmean),.group)) %>% 
  dplyr::select(-c(.group,lower.CL,upper.CL,group)) %>% 
  dplyr::select(
    timing,emmean,CI,n
  ) %>% 
  knitr::kable(
    caption = "Estimated marginal means across N rate timings from a mixed effect model of yield as a function of nitrogen rate, timing and stand.age across 9 site-years. No interactions were detected among nitrogen rate, timing and stand age, but main effects were detected from timing"
  )
```

We reject the Ho that yields were the same regardless of timing. Applying in the fall was estimated to have a higher grain yield than when split in the spring, summer or applied alone in the spring. 

Since the dataset is unbalanced, we reported estimated marginal means, 95% confidence intervals and the number of data points within each timing used in the model.  

### full model, location as fixed effect

We have 4 locations and there is a rationale to model them as fixed effects. This puts a lot of stress on our model by cutting it up by n rate, timing, stand age and location. We end up making a lot of meaningless comparisons and need to reduce the comparisons we make in order to prevent a rank deficient model.

We ran a full factorial model, then would remove interaction terms that were insignificant and rerun the model. 

we removed R100 from the dataset (site years = 8) because it only had one stand age after we removed third stand age for lodging and when stand.age is modelled as a fixed effect the R100 data doesn't provide any utility to testing those hypotheses

```{r}
dat %>% 
  filter(location != "R100") %>%
# removing instances where there was a Fall timing but no N was applied
  filter(timing !="Fall" | fall != 0) %>%
  lmer(
    yield~poly(n.total,2)*timing*stand.age*location + (1|block),.
  ) %>% 
  # summary()
  car::Anova()

# still just stand:location interaction

# reducing model parameters to only significant interactions

dat %>% 
  filter(location != "R100" | stand.age != "3") %>%
# removing instances where there was a Fall timing but no N was applied
  filter(timing !="Fall" | fall != 0) %>%
  filter(location != "R100") %>% 
  lmer(
    yield~poly(n.total,2)+timing+stand.age+
      location+stand.age:location +
      (1|block),.
  ) %>% 
  car::Anova()
```

Change in yield over stand.age was different, Staples when down in year 2 and V17 went up. We will need to separate by location or stand age. 

#### slice by stand age

Slicing by second stand age yielded the only interesting results. 

```{r}
dat %>% 
  filter(location != "R100") %>%
# removing instances where there was a Fall timing but no N was applied
  filter(timing !="Fall" | fall != 0) %>%
  filter(stand.age == "2") %>% 
  lmer(
    yield~poly(n.total,2)*timing*location + (1|block),.
  ) %>% 
  car::Anova()

# look at timing
dat %>% 
  filter(location != "R100") %>%
# removing instances where there was a Fall timing but no N was applied
  filter(timing !="Fall" | fall != 0) %>%
  filter(stand.age == "2") %>% 
  lmer(
    yield~poly(n.total,2)+timing+location + (1|block),.
  ) %>% 
  # summary()
  # car::Anova()
  emmeans(~timing) %>% 
  cld(
    Letters=letters,
    reverse=T
  ) %>% 
  mutate(group = str_trim(.group))-> tuk1

dat %>% 
  filter(location != "R100") %>%
# removing instances where there was a Fall timing but no N was applied
  filter(timing !="Fall" | fall != 0) %>%
  filter(stand.age == "2") %>% 
  group_by(timing) %>% 
  tally() %>% 
  left_join(tuk1) %>% 
  mutate(
    CI = paste0(round(lower.CL),"-",round(upper.CL))
  ) %>% 
  arrange(desc(emmean)) %>% 
  mutate(emmean = paste0(round(emmean),.group)) %>% 
  dplyr::select(-c(.group,lower.CL,upper.CL,group)) %>% 
  dplyr::select(
    timing,emmean,CI,n
  ) %>% 
  knitr::kable(
    caption = "Estimated marginal means across N rate timings from a mixed effect model of yield in second year kernza stands as a function of nitrogen rate, timing and location. No interactions were detected among nitrogen rate, timing and location, but main effects were detected from timing")
```

TAKEAWAY: across 8 site-years, second year yields were higher when N was applied in the fall versus in the spring or a spring summer split. They were also higher in the Fall summer split compared with the spring summer split

#### slice by site

Lastly, we can slice by site and do an independent analysis for each site. This is what Dominic did and I did in the exploratory data analysis.

The main takeaway is we see a great response from staples but not much beyond that. 




