---
title: n-rate timing analysis report
author: jesse puka-beals
date: "`r Sys.Date()`"
# date: "Dec 19, 2022"
output:
  pdf_document
editor_options: 
  chunk_output_type: console
---

```{r, echo=F, message=F}
knitr::opts_chunk$set(echo = F,
                      message = F,
                      include = F, warning = F)

# echo = display code
# message = display code message
# include = include chunk outputs in doc output

library(multcomp)
source("nrate_eda_9dec2022.R")
source("new-dataset.R")

theme_set(theme_bw())

```

# Summary

We need to figure out which data sets to use for comparisons. 

Originally, I wanted to combine datasets to answer different questions, but issues with stand age prevent this

Due to differences across sites, it may make the most sense to analyze sites separately and see if a consistent pattern emerges. This is what Dominic did. 

similar to Dominic, created a new dataframe where N applied is distributed across timings to create a consistent format across sites

\newpage

```{r, include=T}

# this dataset is still not fully tested for accuracy yet. 

dat_new %>% 
  # colnames()
  dplyr::select(location,stand.age, id, fall,spring,summer,yield.kgperha,cumn) %>% 
  mutate(`cumulative N` = fall+spring+summer) %>% 
  head(10) %>% 
  knitr::kable(
    caption = "first 10 rows of new data frame used for analysis"
  )

```

\newpage
```{r, echo=T, include=T, fig.cap="yield response to the combination of fall spring and summer N"}

dat_new  %>% 
  ggplot(aes(fall+spring+summer,yield.kgperha,
             group = stand.age,
             col=stand.age) )+
  geom_point() +
  stat_summary(
    geom="line"
  ) +
  facet_wrap(~location)
```

\newpage

```{r,include=T, fig.cap="yield response to fertilizer within timing"}
dat_new  %>% 
  ggplot(aes(fall,yield.kgperha,
             group = stand.age,
             col=stand.age) )+
  geom_point() +
  stat_summary(
    geom="line"
  ) +
  facet_wrap(~location)

dat_new  %>% 
  ggplot(aes(spring,yield.kgperha,
             group = stand.age,
             col=stand.age) )+
  geom_point() +
  stat_summary(
    geom="line"
  ) +
  facet_wrap(~location)

dat_new  %>% 
  ggplot(aes(summer,yield.kgperha,
             group = stand.age,
             col=stand.age) )+
  geom_point() +
  stat_summary(
    geom="line"
  ) +
  facet_wrap(~location)
```

\newpage

```{r}

dat_new %>% 
  pivot_longer(
    c(fall,spring,summer)
  ) %>% 
  # View()
  ggplot(aes(value,yield.kgperha,
             color=name,
             shape = stand.age)) +
  geom_point() +
  facet_wrap(~location)
```

# Methods

Yield data was collected from 3 sites. All sites have control plots. 

V17: treatments initiated in spring of year 1, 3 years of yield data. No summer N, fall (80N), spring (80N) or split (40N spring, 40N fall)

R100: treatments initiated in spring of year 2, 2 years of yield data. Spring and fall N at 60N, summer N varies. 

staples: treatments initiated in spring of year 1, 3 years of yield data. Spring and fall N at 60N, summer N varies. 

Sites are confounded by planting year, stand.age, soil type, row spacing. 

Fall treatments only begin to occur in year 2. 

\newpage

```{r, include=T}
dat_v17 %>% 
  bind_rows(dat_staples) %>% 
  bind_rows(dat_r100) %>% 
  # colnames()
  # group_by(location,year,stand.age) %>%
  group_by(year, location, stand.age) %>% 
  # glimpse()
  mutate(year = fct_relevel(year, 
                            "2018", 
                            .before = "2019")) %>% 
  summarise(n=n()) %>% 
  knitr::kable(
    caption = "year X location X stand.age"
  )
```


```{r, include=T, echo=F}


dat_v17 %>% 
  bind_rows(dat_r100) %>% 
  bind_rows(dat_staples) %>% 
  group_by(location, stand.age) %>% 
  summarise(n=n()) %>% 
  knitr::kable(
    caption = "location X stand age"
  )

```

\newpage

```{r, include=T, echo=F}
source("nrate_eda_9dec2022.R")

dat_v17 %>% 
  bind_rows(dat_r100) %>% 
  bind_rows(dat_staples) %>% 
  group_by(location, treatment) %>% 
  summarise(n=n()) %>% 
  knitr::kable(
    caption = "location X treatment"
  )
  

```


\newpage

## treatment table

+------------+------------+------------+----------------------------+
| location   | harvest    | timing[^1] | N applied prior to harvest |
+============+============+============+============================+
| V17        | 1          | fall       | no                         |
+------------+------------+------------+----------------------------+
| V17        | 1          | spring     | yes                        |
+------------+------------+------------+----------------------------+
| V17        | 1          | summer     | no                         |
+------------+------------+------------+----------------------------+
| V17        | 2          | fall       | yes                        |
+------------+------------+------------+----------------------------+
| V17        | 2          | spring     | yes                        |
+------------+------------+------------+----------------------------+
| V17        | 2          | summer     | no                         |
+------------+------------+------------+----------------------------+
| V17        | 3          | fall       | yes                        |
+------------+------------+------------+----------------------------+
| V17        | 3          | spring     | yes                        |
+------------+------------+------------+----------------------------+
| V17        | 3          | summer     | no                         |
+------------+------------+------------+----------------------------+
| Staples    | 1          | fall       | no                         |
+------------+------------+------------+----------------------------+
| Staples    | 1          | spring     | yes                        |
+------------+------------+------------+----------------------------+
| Staples    | 1          | summer     | yes                        |
+------------+------------+------------+----------------------------+
| Staples    | 2          | fall       | yes                        |
+------------+------------+------------+----------------------------+
| Staples    | 2          | spring     | yes                        |
+------------+------------+------------+----------------------------+
| Staples    | 2          | summer     | yes                        |
+------------+------------+------------+----------------------------+
| Staples    | 3          | fall       | yes                        |
+------------+------------+------------+----------------------------+
| Staples    | 3          | spring     | yes                        |
+------------+------------+------------+----------------------------+
| Staples    | 3          | summer     | yes                        |
+------------+------------+------------+----------------------------+
| R100       | 1          | fall       | no                         |
+------------+------------+------------+----------------------------+
| R100       | 1          | spring     | yes                        |
+------------+------------+------------+----------------------------+
| R100       | 1          | summer     | yes                        |
+------------+------------+------------+----------------------------+
| R100       | 2          | fall       | yes                        |
+------------+------------+------------+----------------------------+
| R100       | 2          | spring     | yes                        |
+------------+------------+------------+----------------------------+
| R100       | 2          | summer     | yes                        |
+------------+------------+------------+----------------------------+
| R100       | 3          | fall       | yes                        |
+------------+------------+------------+----------------------------+
| R100       | 3          | spring     | yes                        |
+------------+------------+------------+----------------------------+
| R100       | 3          | summer     | yes                        |
+------------+------------+------------+----------------------------+

[^1]: fall timing is applied in year prior to harvest

At V17, plots with a spring timing got fertilized prior to the first kernza harvest. If any of that N remains in the soil and rolls over into the next year, it will make spring N plots look higher yielding then they are.

At R100 and Staples, this effect will be diluted by the summer applied N

\newpage

# Datasets

```{r, include=F}

dat_staples %>% 
  # group_by(stand.age) %>% 
  group_by(treatment) %>% 
  summarise(n=n())
dat_r100%>% 
  # group_by(stand.age) %>% 
  group_by(treatment) %>% 
  summarise(n=n())
dat_v17 %>% 
  # group_by(stand.age) %>% 
  group_by(treatment) %>% 
  summarise(n=n())

```

+----------+------------+----------+---------------------------------------+---------+
| location | stand.ages | n        | treatments                            | n       |
+==========+============+==========+=======================================+=========+
| Staples  | 1,2,3      | 54,54,54 | n=18, same as r100                    | 9       |
+----------+------------+----------+---------------------------------------+---------+
| R100     | 2,3        | 54,54    | n=18; same as staples                 | 6       |
+----------+------------+----------+---------------------------------------+---------+
| V17      | 1,2,3      | 16,16,16 | control, 80 fall, 80 split, 80 spring | 12      |
+----------+------------+----------+---------------------------------------+---------+

+-------------------+-------------+--------------------+------------+---------------------------------+----------------------------------------+
| balanced datasets | name        | locations          | stand.ages | treatments                      | model                                  |
+===================+=============+====================+============+=================================+========================================+
| timing            | dat_timing  | Staples, V17       | 1,2,3      | control, spring only            | y\~timing X location X stand.age       |
+-------------------+-------------+--------------------+------------+---------------------------------+----------------------------------------+
| timing2           | dat_timing2 | Staples, V17, R100 | 2,3        | control, spring only, fall only | y\~timing X location X stand.age       |
+-------------------+-------------+--------------------+------------+---------------------------------+----------------------------------------+
| nrate             | dat_nrate   | Staples            | 1,2,3      | all                             | y\~updatednrate X location X stand.age |
+-------------------+-------------+--------------------+------------+---------------------------------+----------------------------------------+
| nrate older stand | dat_old     | Staples, R100      | 2,3        | all                             | y\~treatment X location X stand.age    |
+-------------------+-------------+--------------------+------------+---------------------------------+----------------------------------------+

+-------------------+----------------------------------------------------------------------------------------------+
| balanced datasets | testable hypotheses                                                                          |
+===================+==============================================================================================+
| dat_timing        | Yields are similar regardless of fertilization                                               |
|                   |                                                                                              |
|                   | Yields are similar regardless of timing (treatment != control)                               |
|                   |                                                                                              |
|                   | Yields are similar regardless of stand.age or location                                       |
+-------------------+----------------------------------------------------------------------------------------------+
| dat_nrate         | yields are similar regardless of fertilization                                               |
|                   |                                                                                              |
|                   | yields are similar regardless of timing\*rate (treatment != control)                         |
|                   |                                                                                              |
|                   | Yields are similar regardless of stand.age or location                                       |
|                   |                                                                                              |
|                   | Yields are similar regardless of the amount of N applied prior to harvest                    |
|                   |                                                                                              |
|                   | Yield response to N is best modeled as no effect vs a linear, quadratic and quadratic planar |
+-------------------+----------------------------------------------------------------------------------------------+
| dat_old           | Yields in older stands are similar regardless of fertilization                               |
|                   |                                                                                              |
|                   | yields in older stands are similar regardless of timing (treatment != control)               |
|                   |                                                                                              |
|                   | yields in older stands are similar regardless of timing\*rate (treatment != control)         |
|                   |                                                                                              |
|                   | yields in older stands are similar regardless of the amount of N applied prior to harvest    |
|                   |                                                                                              |
|                   | yields are similar regardless of stand age and location                                      |
+-------------------+----------------------------------------------------------------------------------------------+

**Additional questions/framings**

What nitrogen rate results in highest grain yield?

As stand age increases, does optimal nitrogen rate increase?

**Dominic's findings**

R100: timing does not matter, 0-60N maximizes yield

Stapes: applying N in fall near 80 lb A is best

V17: fall application is best (including split)

**Jesse's findings**

R100: does not make a lot of sense

\newpage

# Yield estimate issue

This issue was caused by converting numeric yield to a factor category and then converting it back to numeric. Very strange, but the problem is resolved

\newpage


# Analysis

```{r}

dat_v17_tiny %>%
  bind_rows(dat_staples_tiny) -> dat_timing

dat_staples -> dat_nrate

dat_r100 %>%
  bind_rows(dat_staples) %>%
  filter(stand.age != "1") -> dat_old

```

## Dat_timing

```{r no fall}

dat_timing %>%
  filter(timing != "fall") %>%
  lm(yield.kgperha~timing*stand.age*location,
     .
     ) %>%
  # anova()
  emmeans(~timing)

```

We reject Ho that yields are similar between plots that were and were no fertilized in the spring across all 3 years at V17 and R100.

```{r fall and spring but only years 2 and 3}

# we need to add in R100 now

dat_timing %>%
  filter(stand.age != "1") %>%
  lm(yield.kgperha~timing*stand.age*location,
     .
     ) %>%
  anova()
  # emmeans(~timing)


```

```{r}
## Mixed effect models don't work here
# Block explains nothing and locations are too few to treat as random

# lmer(
#   yield.kgperha~timing*stand.age+
#     (1|location/block),
#   dat_timing
# ) %>%
#   car::Anova()


dat_timing %>%
  # distinct(location)
  # filter(location=="Staples") %>%
  filter(location=="RSMT V17") %>%
  lm(
    # yield.kgperha~timing*location*stand.age,
    yield.kgperha~timing*stand.age,
    .
  ) %>%
  # summary()
  anova()
  # emmeans(~timing*stand.age) %>%
  # cld(Letters=letters, reverse=T)

# 3-way interaction with location and stand.age
# staples interaction between timing and stand age

# interaction
dat_timing %>%
  filter(location=="Staples") %>%
  # filter(location=="RSMT V17") %>%
  ggplot(aes(timing,yield.kgperha,
             group = stand.age,
             col=stand.age)) +
  stat_summary(geom = "line")

# This interaction is due to year1 fall essentially being a control

dat_timing %>%
  filter(stand.age != "1" | timing != "fall") %>%
  filter(location == "Staples") %>%
  # ggplot(aes(timing,yield.kgperha,
  #            group = stand.age,
  #            color=stand.age)) +
  # stat_summary(geom = "line")
  lm(
    yield.kgperha~timing*stand.age,
    .
  ) %>%
  anova()


```

+--------------+--------------+---------------------------+--------------+
| dataset      | location     | fixed effects             | p-value      |
+==============+==============+===========================+==============+
| dat_timing   | combined     | timing:location:stand.age | \*\*         |
+--------------+--------------+---------------------------+--------------+
| dat_timing   | staples      | timing                    | \*\*         |
+--------------+--------------+---------------------------+--------------+
| dat_timing   | staples      | stand.age                 | \*\*\*       |
+--------------+--------------+---------------------------+--------------+
| dat_timing   | staples      | timing\*stand.age         | \*           |
+--------------+--------------+---------------------------+--------------+
| dat_timing   | v17          | timing                    | \*\*\*       |
+--------------+--------------+---------------------------+--------------+
| dat_timing   | v17          | stand.age                 | \*\*\*       |
+--------------+--------------+---------------------------+--------------+
| dat_timing   | v17          | timing\*stand.age         | NS           |
+--------------+--------------+---------------------------+--------------+

The effects of timing and stand age were not consistent across sites, and required each site be analyzed separately.

*Yields are similar regardless of fertilization*

Staples: there is an interaction between the effect of fertilization and stand.age.

R100:

Yields are similar regardless of timing (treatment != control)

Yields are similar regardless of stand.age or location
