---
title: n-rate timing report
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---

```{r, echo=F, message=F}
knitr::opts_chunk$set(echo = F,message = F, include = F)

# echo = display code
# message = display code message
# include = include chunk outputs in doc output

library(multcomp)

```


# Questions

Were the fall treatments applied right after planting or the following year?\
Are there plots that received nitrogen in the fall prior to their first harvest?

Based on data, it appears the plots with a fall treatment applied N in fall for Staples. This complicates how we analyze timing and the reason we have in interaction at staples

We may have to make a new dataset--remove the fall treatments because they are offset in their fertilization...this is why we need to do a better job thinking out how the data will be analyzed. Doing this stuff weirdly after the fact is annoying because it is heartbreaking getting rid of data. This was just not planned well, and we shouldn't have taken the data. But also, it's fine. I always go back and forth



# Datasets

```{r, include=F}

source("nrate_eda_9dec2022.R")
dat_staples %>% 
  # group_by(stand.age) %>% 
  group_by(treatment) %>% 
  summarise(n=n())
dat_r100%>% 
  # group_by(stand.age) %>% 
  group_by(treatment) %>% 
  summarise(n=n())
dat_v17 %>% 
  # group_by(stand.age) %>% 
  group_by(treatment) %>% 
  summarise(n=n())

```

+----------+------------+----------+---------------------------------------+---------+
| location | stand.ages | n        | treatments                            | n       |
+==========+============+==========+=======================================+=========+
| Staples  | 1,2,3      | 54,54,54 | n=18, same as r100                    | 9       |
+----------+------------+----------+---------------------------------------+---------+
| R100     | 2,3        | 54,54    | n=18; same as staples                 | 6       |
+----------+------------+----------+---------------------------------------+---------+
| V17      | 1,2,3      | 16,16,16 | control, 80 fall, 80 split, 80 spring | 12      |
+----------+------------+----------+---------------------------------------+---------+

+---------------------------+------------+---------------+------------+---------------------------------+---------------------------------+
| balanced datasets         |            | locations     | stand.ages | treatments                      | model                           |
+===========================+============+===============+============+=================================+=================================+
| timing dataset            | dat_timing | Staples, V17  | 1,2,3      | control, spring only, fall only | y\~treatmentXlocationXstand.age |
+---------------------------+------------+---------------+------------+---------------------------------+---------------------------------+
| nrate dataset             | dat_nrate  | Staples       | 1,2,3      | all                             | y\~treatmentXlocationXstand.age |
+---------------------------+------------+---------------+------------+---------------------------------+---------------------------------+
| nrate older stand dataset | dat_old    | Staples, R100 | 2,3        | all                             | y\~treatmentXlocationXstand.age |
+---------------------------+------------+---------------+------------+---------------------------------+---------------------------------+

+-------------------+----------------------------------------------------------------------------------------------+
| balanced datasets | testable hypotheses                                                                          |
+===================+==============================================================================================+
| dat_timing        | Yields are similar regardless of fertilization                                               |
|                   |                                                                                              |
|                   | Yields are similar regardless of timing (treatment != control)                               |
|                   |                                                                                              |
|                   | Yields are similar regardless of stand.age or location                                       |
+-------------------+----------------------------------------------------------------------------------------------+
| dat_nrate         | yields are similar regardless of fertilization                                               |
|                   |                                                                                              |
|                   | yields are similar regardless of timing\*rate (treatment != control)                         |
|                   |                                                                                              |
|                   | Yields are similar regardless of stand.age or location                                       |
|                   |                                                                                              |
|                   | Yields are similar regardless of the amount of N applied prior to harvest                    |
|                   |                                                                                              |
|                   | Yield response to N is best modeled as no effect vs a linear, quadratic and quadratic planar |
+-------------------+----------------------------------------------------------------------------------------------+
| dat_old           | Yields in older stands are similar regardless of fertilization                               |
|                   |                                                                                              |
|                   | yields in older stands are similar regardless of timing (treatment != control)               |
|                   |                                                                                              |
|                   | yields in older stands are similar regardless of timing\*rate (treatment != control)         |
|                   |                                                                                              |
|                   | yields in older stands are similar regardless of the amount of N applied prior to harvest    |
|                   |                                                                                              |
|                   | yields are similar regardless of stand age and location                                      |
+-------------------+----------------------------------------------------------------------------------------------+

**Additional questions/framings**

What nitrogen rate results in highest grain yield?

As stand age increases, does optimal nitrogen rate increase?

**Dominic's findings**

R100: timing does not matter, 0-60N maximizes yield

Stapes: applying N in fall near 80 lb A is best

V17: fall application is best (including split)

**Jesse's findings**

R100: does not make a lot of sense

# Analysis

```{r}

dat_v17_tiny %>% 
  bind_rows(dat_staples_tiny) -> dat_timing

dat_staples -> dat_nrate

dat_r100 %>% 
  bind_rows(dat_staples) %>% 
  filter(stand.age != "1") -> dat_old

```

## Dat_timing

```{r}
## Mixed effect models don't work here
# Block explains nothing and locations are too few to treat as random
  
# lmer(
#   yield.kgperha~timing*stand.age+
#     (1|location/block),
#   dat_timing
# ) %>% 
#   car::Anova()


# dat_timing %>% 
  # distinct(location)
  lm(
    # yield.kgperha~timing*location*stand.age,
    yield.kgperha~timing*stand.age,
    filter(dat_timing,location=="Staples")
    # filter(dat_timing,location=="RSMT V17")
    # .
  ) %>% 
  # summary()
  # anova()
  emmeans(~timing*stand.age) %>% 
  cld(Letters=letters, reverse=T)

# 3-way interaction with location and stand.age
# staples interaction between timing and stand age
  
# interaction
dat_timing %>% 
  filter(location=="Staples") %>%
  # filter(location=="RSMT V17") %>% 
  ggplot(aes(timing,yield.kgperha,
             group = stand.age,
             col=stand.age)) +
  stat_summary(geom = "line") 

# This interaction is due to year1 fall essentially being a control

dat_timing %>% 
  filter(stand.age != "1" | timing != "fall") %>% 
  filter(location == "Staples") %>% 
  # ggplot(aes(timing,yield.kgperha,
  #            group = stand.age,
  #            color=stand.age)) +
  # stat_summary(geom = "line")
  lm(
    yield.kgperha~timing*stand.age,
    .
  ) %>% 
  anova()


```

+-------------+-------------+---------------------------+------------------+
| dataset     | location    | fixed effects             | p-value          |
+=============+=============+===========================+==================+
| dat_timing  | combined    | timing:location:stand.age | \*\*             |
+-------------+-------------+---------------------------+------------------+
| dat_timing  | staples     | timing                    | \*\*             |
+-------------+-------------+---------------------------+------------------+
| dat_timing  | staples     | stand.age                 | \*\*\*           |
+-------------+-------------+---------------------------+------------------+
| dat_timing  | staples     | timing\*stand.age         | \*               |
+-------------+-------------+---------------------------+------------------+
| dat_timing  | v17         | timing                    | \*\*\*           |
+-------------+-------------+---------------------------+------------------+
| dat_timing  | v17         | stand.age                 | \*\*\*           |
+-------------+-------------+---------------------------+------------------+
| dat_timing  | v17         | timing\*stand.age         |                  |
+-------------+-------------+---------------------------+------------------+

The effects of timing and stand age were not consistent across sites, and required each site be analyzed separately.

*Yields are similar regardless of fertilization*

Staples: there is an interaction between the effect of fertilization and stand.age.

R100:

Yields are similar regardless of timing (treatment != control)

Yields are similar regardless of stand.age or location
