---
title: n-rate timing analysis report
author: jesse puka-beals
date: "`r Sys.Date()`"
output:
  pdf_document
editor_options: 
  chunk_output_type: console
---

```{r, echo=F, message=F}
knitr::opts_chunk$set(echo = F,
                      message = F,
                      include = F, warning = F)

# echo = display code
# message = display code message
# include = include chunk outputs in doc output

library(multcomp)
source("nrate_eda_9dec2022.R")
source("new-dataset.R")

theme_set(theme_bw())

```

# Introduction

Sites are all different with treatments. Staples makes most sense.

How does cumulative N impact yield?

Does timing of N change this impact?

What might've caused differences among sites?

# Expected Findings

Yields should be ~500 kg ha, going down in third year

Yields should be higher in plots receiving more N

No idea if timing matters and no idea if it's consistent


\newpage

# Methods

Yield data was collected from 3 sites. All sites have control plots.

V17: treatments initiated in spring of year 1, 3 years of yield data. No summer N, fall (80N), spring (80N) or split (40N spring, 40N fall)

R100: treatments initiated in spring of year 2, 2 years of yield data. Spring and fall N at 60N, summer N varies.

staples: treatments initiated in spring of year 1, 3 years of yield data. Spring and fall N at 60N, summer N varies.

Sites are confounded by planting year, stand.age, soil type, row spacing.

Fall treatments only begin to occur in year 2.

\newpage

```{r, include=T}
dat_v17 %>% 
  bind_rows(dat_staples) %>% 
  bind_rows(dat_r100) %>% 
  # colnames()
  # group_by(location,year,stand.age) %>%
  group_by(year, location, stand.age) %>% 
  # glimpse()
  mutate(year = fct_relevel(year, 
                            "2018", 
                            .before = "2019")) %>% 
  summarise(n=n()) %>% 
  knitr::kable(
    caption = "year X location X stand.age"
  )
```

```{r, include=T, echo=F}


dat_v17 %>% 
  bind_rows(dat_r100) %>% 
  bind_rows(dat_staples) %>% 
  group_by(location, stand.age) %>% 
  summarise(n=n()) %>% 
  knitr::kable(
    caption = "location X stand age"
  )

```

\newpage

```{r, include=T, echo=F}
source("nrate_eda_9dec2022.R")

dat_v17 %>% 
  bind_rows(dat_r100) %>% 
  bind_rows(dat_staples) %>% 
  group_by(location, treatment) %>% 
  summarise(n=n()) %>% 
  knitr::kable(
    caption = "location X treatment"
  )
  

```

\newpage

## treatment table

+-------------+-------------+-------------+----------------------------+
| location    | harvest     | timing[^1]  | N applied prior to harvest |
+=============+=============+=============+============================+
| V17         | 1           | fall        | no                         |
+-------------+-------------+-------------+----------------------------+
| V17         | 1           | spring      | yes                        |
+-------------+-------------+-------------+----------------------------+
| V17         | 1           | summer      | no                         |
+-------------+-------------+-------------+----------------------------+
| V17         | 2           | fall        | yes                        |
+-------------+-------------+-------------+----------------------------+
| V17         | 2           | spring      | yes                        |
+-------------+-------------+-------------+----------------------------+
| V17         | 2           | summer      | no                         |
+-------------+-------------+-------------+----------------------------+
| V17         | 3           | fall        | yes                        |
+-------------+-------------+-------------+----------------------------+
| V17         | 3           | spring      | yes                        |
+-------------+-------------+-------------+----------------------------+
| V17         | 3           | summer      | no                         |
+-------------+-------------+-------------+----------------------------+
| Staples     | 1           | fall        | no                         |
+-------------+-------------+-------------+----------------------------+
| Staples     | 1           | spring      | yes                        |
+-------------+-------------+-------------+----------------------------+
| Staples     | 1           | summer      | yes                        |
+-------------+-------------+-------------+----------------------------+
| Staples     | 2           | fall        | yes                        |
+-------------+-------------+-------------+----------------------------+
| Staples     | 2           | spring      | yes                        |
+-------------+-------------+-------------+----------------------------+
| Staples     | 2           | summer      | yes                        |
+-------------+-------------+-------------+----------------------------+
| Staples     | 3           | fall        | yes                        |
+-------------+-------------+-------------+----------------------------+
| Staples     | 3           | spring      | yes                        |
+-------------+-------------+-------------+----------------------------+
| Staples     | 3           | summer      | yes                        |
+-------------+-------------+-------------+----------------------------+
| R100        | 1           | fall        | no                         |
+-------------+-------------+-------------+----------------------------+
| R100        | 1           | spring      | yes                        |
+-------------+-------------+-------------+----------------------------+
| R100        | 1           | summer      | yes                        |
+-------------+-------------+-------------+----------------------------+
| R100        | 2           | fall        | yes                        |
+-------------+-------------+-------------+----------------------------+
| R100        | 2           | spring      | yes                        |
+-------------+-------------+-------------+----------------------------+
| R100        | 2           | summer      | yes                        |
+-------------+-------------+-------------+----------------------------+
| R100        | 3           | fall        | yes                        |
+-------------+-------------+-------------+----------------------------+
| R100        | 3           | spring      | yes                        |
+-------------+-------------+-------------+----------------------------+
| R100        | 3           | summer      | yes                        |
+-------------+-------------+-------------+----------------------------+

[^1]: fall timing is applied in year prior to harvest

At V17, plots with a spring timing got fertilized prior to the first kernza harvest. If any of that N remains in the soil and rolls over into the next year, it will make spring N plots look higher yielding then they are.

At R100 and Staples, this effect will be diluted by the summer applied N

\newpage

# Datasets


```{r, dataset}

dat_new %>%
  mutate(across(starts_with("height"),
                as.numeric)) %>%
  mutate(height = (height_1 + height_2 + height_3 + height_4 + height_5) /
           5) %>%
  dplyr::select(location,
                stand.age,
                id,
                fall,
                spring,
                summer,
                yield.kgperha,
                lodging,
                height) %>%
  mutate(cumn = fall + spring + summer,
         .before = yield.kgperha) %>%
  mutate(lodging = as.numeric(lodging)) -> dat_new2

dat_new2 %>% 
  glimpse()
```



```{r, include=F}

dat_staples %>% 
  # group_by(stand.age) %>% 
  group_by(treatment) %>% 
  summarise(n=n())
dat_r100%>% 
  # group_by(stand.age) %>% 
  group_by(treatment) %>% 
  summarise(n=n())
dat_v17 %>% 
  # group_by(stand.age) %>% 
  group_by(treatment) %>% 
  summarise(n=n())

```

+----------+------------+----------+---------------------------------------+---------+
| location | stand.ages | n        | treatments                            | n       |
+==========+============+==========+=======================================+=========+
| Staples  | 1,2,3      | 54,54,54 | n=18, same as r100                    | 9       |
+----------+------------+----------+---------------------------------------+---------+
| R100     | 2,3        | 54,54    | n=18; same as staples                 | 6       |
+----------+------------+----------+---------------------------------------+---------+
| V17      | 1,2,3      | 16,16,16 | control, 80 fall, 80 split, 80 spring | 12      |
+----------+------------+----------+---------------------------------------+---------+

+-------------------+-------------+--------------------+------------+---------------------------------+----------------------------------------+
| balanced datasets | name        | locations          | stand.ages | treatments                      | model                                  |
+===================+=============+====================+============+=================================+========================================+
| timing            | dat_timing  | Staples, V17       | 1,2,3      | control, spring only            | y\~timing X location X stand.age       |
+-------------------+-------------+--------------------+------------+---------------------------------+----------------------------------------+
| timing2           | dat_timing2 | Staples, V17, R100 | 2,3        | control, spring only, fall only | y\~timing X location X stand.age       |
+-------------------+-------------+--------------------+------------+---------------------------------+----------------------------------------+
| nrate             | dat_nrate   | Staples            | 1,2,3      | all                             | y\~updatednrate X location X stand.age |
+-------------------+-------------+--------------------+------------+---------------------------------+----------------------------------------+
| nrate older stand | dat_old     | Staples, R100      | 2,3        | all                             | y\~treatment X location X stand.age    |
+-------------------+-------------+--------------------+------------+---------------------------------+----------------------------------------+

+-------------------+----------------------------------------------------------------------------------------------+
| balanced datasets | testable hypotheses                                                                          |
+===================+==============================================================================================+
| dat_timing        | Yields are similar regardless of fertilization                                               |
|                   |                                                                                              |
|                   | Yields are similar regardless of timing (treatment != control)                               |
|                   |                                                                                              |
|                   | Yields are similar regardless of stand.age or location                                       |
+-------------------+----------------------------------------------------------------------------------------------+
| dat_nrate         | yields are similar regardless of fertilization                                               |
|                   |                                                                                              |
|                   | yields are similar regardless of timing\*rate (treatment != control)                         |
|                   |                                                                                              |
|                   | Yields are similar regardless of stand.age or location                                       |
|                   |                                                                                              |
|                   | Yields are similar regardless of the amount of N applied prior to harvest                    |
|                   |                                                                                              |
|                   | Yield response to N is best modeled as no effect vs a linear, quadratic and quadratic planar |
+-------------------+----------------------------------------------------------------------------------------------+
| dat_old           | Yields in older stands are similar regardless of fertilization                               |
|                   |                                                                                              |
|                   | yields in older stands are similar regardless of timing (treatment != control)               |
|                   |                                                                                              |
|                   | yields in older stands are similar regardless of timing\*rate (treatment != control)         |
|                   |                                                                                              |
|                   | yields in older stands are similar regardless of the amount of N applied prior to harvest    |
|                   |                                                                                              |
|                   | yields are similar regardless of stand age and location                                      |
+-------------------+----------------------------------------------------------------------------------------------+

**Additional questions/framings**

What nitrogen rate results in highest grain yield?

As stand age increases, does optimal nitrogen rate increase?

**Dominic's findings**

R100: timing does not matter, 0-60N maximizes yield

Stapes: applying N in fall near 80 lb A is best

V17: fall application is best (including split)

**Jesse's findings**

R100: does not make a lot of sense

\newpage

# Yield

## Yield

### with lodging considered

\newpage

```{r, include=T, fig.cap="Cumulative yield response to cumulative N application"}

dat_new %>% 
  mutate(lodging.num = as.numeric(lodging)) %>% 
  mutate(lodging.num = replace_na(lodging.num,1)) %>% 
  mutate(lodging_status = if_else(
    lodging.num > 5, 
    "yes",
    "no / unknown"
  )) %>% 
  filter(lodging_status == "no / unknown") %>% 
  dplyr::select(location,stand.age, id, fall,spring,summer,yield.kgperha,lodging) %>% 
  mutate(cumn_priortoharvest = fall+spring+summer) %>% 
  group_by(location,id) %>% 
  # summarise(cumn_total = fall+spring+summer)
  # summarise(cumn_total = cumsum(fall+spring+summer))
  summarise(cumn_total=sum(fall+spring+fall),
            cumyield_total = sum(yield.kgperha
                                 # na.rm = T #V17 101 has missing yield in year2
                                 ))  %>% 
  # knitr::kable()
  ggplot(aes(
    cumn_total,
    cumyield_total,
    col=location
  )) +
  stat_summary(geom="line") +
  labs(caption = "R100 shows no N response with and without lodging scores > 5 in dataset
       this shows dataset where lodging greater than 5 was filtered out")
```




## Cumulative yield

\newpage

```{r, include=T, fig.cap="Cumulative yield response to cumulative N application"}

dat_new %>% 
  mutate(lodging.num = as.numeric(lodging)) %>% 
  mutate(lodging.num = replace_na(lodging.num,1)) %>% 
  mutate(lodging_status = if_else(
    lodging.num > 5, 
    "yes",
    "no / unknown"
  )) %>% 
  filter(lodging_status == "no / unknown") %>% 
  dplyr::select(location,stand.age, id, fall,spring,summer,yield.kgperha,lodging) %>% 
  mutate(cumn_priortoharvest = fall+spring+summer) %>% 
  group_by(location,id) %>% 
  # summarise(cumn_total = fall+spring+summer)
  # summarise(cumn_total = cumsum(fall+spring+summer))
  summarise(cumn_total=sum(fall+spring+fall),
            cumyield_total = sum(yield.kgperha
                                 # na.rm = T #V17 101 has missing yield in year2
                                 ))  %>% 
  # knitr::kable()
  ggplot(aes(
    cumn_total,
    cumyield_total,
    col=location
  )) +
  stat_summary(geom="line") +
  labs(caption = "R100 shows no N response with and without lodging scores > 5 in dataset
       this shows dataset where lodging greater than 5 was filtered out")
```



\newpage

# Analysis

```{r}

dat_v17_tiny %>%
  bind_rows(dat_staples_tiny) -> dat_timing

dat_staples -> dat_nrate

dat_r100 %>%
  bind_rows(dat_staples) %>%
  filter(stand.age != "1") -> dat_old

```

## Dat_timing

```{r no fall}

dat_timing %>%
  filter(timing != "fall") %>%
  lm(yield.kgperha~timing*stand.age*location,
     .
     ) %>%
  # anova()
  emmeans(~timing)

```

We reject Ho that yields are similar between plots that were and were no fertilized in the spring across all 3 years at V17 and R100.

```{r fall and spring but only years 2 and 3}

# we need to add in R100 now

dat_timing %>%
  filter(stand.age != "1") %>%
  lm(yield.kgperha~timing*stand.age*location,
     .
     ) %>%
  anova()
  # emmeans(~timing)


```

```{r}
## Mixed effect models don't work here
# Block explains nothing and locations are too few to treat as random

# lmer(
#   yield.kgperha~timing*stand.age+
#     (1|location/block),
#   dat_timing
# ) %>%
#   car::Anova()


dat_timing %>%
  # distinct(location)
  # filter(location=="Staples") %>%
  filter(location=="RSMT V17") %>%
  lm(
    # yield.kgperha~timing*location*stand.age,
    yield.kgperha~timing*stand.age,
    .
  ) %>%
  # summary()
  anova()
  # emmeans(~timing*stand.age) %>%
  # cld(Letters=letters, reverse=T)

# 3-way interaction with location and stand.age
# staples interaction between timing and stand age

# interaction
dat_timing %>%
  filter(location=="Staples") %>%
  # filter(location=="RSMT V17") %>%
  ggplot(aes(timing,yield.kgperha,
             group = stand.age,
             col=stand.age)) +
  stat_summary(geom = "line")

# This interaction is due to year1 fall essentially being a control

dat_timing %>%
  filter(stand.age != "1" | timing != "fall") %>%
  filter(location == "Staples") %>%
  # ggplot(aes(timing,yield.kgperha,
  #            group = stand.age,
  #            color=stand.age)) +
  # stat_summary(geom = "line")
  lm(
    yield.kgperha~timing*stand.age,
    .
  ) %>%
  anova()


```

+--------------+--------------+---------------------------+--------------+
| dataset      | location     | fixed effects             | p-value      |
+==============+==============+===========================+==============+
| dat_timing   | combined     | timing:location:stand.age | \*\*         |
+--------------+--------------+---------------------------+--------------+
| dat_timing   | staples      | timing                    | \*\*         |
+--------------+--------------+---------------------------+--------------+
| dat_timing   | staples      | stand.age                 | \*\*\*       |
+--------------+--------------+---------------------------+--------------+
| dat_timing   | staples      | timing\*stand.age         | \*           |
+--------------+--------------+---------------------------+--------------+
| dat_timing   | v17          | timing                    | \*\*\*       |
+--------------+--------------+---------------------------+--------------+
| dat_timing   | v17          | stand.age                 | \*\*\*       |
+--------------+--------------+---------------------------+--------------+
| dat_timing   | v17          | timing\*stand.age         | NS           |
+--------------+--------------+---------------------------+--------------+

The effects of timing and stand age were not consistent across sites, and required each site be analyzed separately.

*Yields are similar regardless of fertilization*

Staples: there is an interaction between the effect of fertilization and stand.age.

R100:

Yields are similar regardless of timing (treatment != control)

Yields are similar regardless of stand.age or location


