---
title: "Rosemount R-100"
author: "Dominic Christensen"
date: "06/17/2022"
output: html_document
---

# Rosemount R-100 tables, figures, and statistics

# Set working directory, load packages, read data files, and modify character codes
```{r}
# Load packages
library(readxl)
library(tidyverse)
library(emmeans)
library(lme4)
library(lattice)
library(patchwork)
library(car)
library(ggeffects)


# Read data file
r100<- read.csv("Rosemount_R100.csv")

# Modify character codes
str(r100)
r100$stand.age<- as.factor(r100$stand.age)
r100$Block<- as.factor(r100$Block)

# Note: One modification occured from original master file and that was changing "Control" to "control".

```


```{r jesse exploring}

r100 %>% 
  ggplot(aes(yield.kgperha)) +
  stat_bin() +
  labs(caption = "if we filter values below 150, it would be normally distributed\nthis could be due to lodging")

r100 %>% 
  ggplot(aes(yield.kgperha)) +
  stat_bin(aes(fill=stand.age)) +
  labs(caption = "nevermind, explained by stand age")

r100 %>% 
  ggplot(aes(yield.kgperha)) +
  # geom_density(aes(col=Ntiming)) +
  stat_bin(aes(fill=Ntiming)) +
  facet_wrap(~stand.age) +
  labs(caption = "year3 looks weird, could be lodging")

r100 %>% 
  ggplot(aes(yield.kgperha)) +
  geom_density(aes(col=factor(Napplied_update))) +
  facet_wrap(~stand.age) +
  labs(caption = "is lodging explaining why low rates are normally distributed but others are skewed?")

# Just start with stand age = 2, that's normal
# stand age 3 might need to be modeled with a glm
```

```{r jesse exploring more}

lme.jpb.1 <- lmer(yield.kgperha~Napplied_update*Ntiming+(1|Block), data=subset(r100,stand.age=="2"))
summary(lme.jpb.1)
# this model is too complex for data, rank deficient

lme.jpb.2 <- lmer(yield.kgperha~(Napplied_update)*Ntiming+(1|Block), data=subset(r100,stand.age=="2"))
summary(lme.jpb.2)
Anova(lme.jpb.2)

r100 %>% 
  # filter(stand.age=="2") %>%
  ggplot(aes(Napplied_update,yield.kgperha)) +
  geom_point(aes(col=Ntiming)) +
  geom_smooth(method="lm") +
  geom_smooth(method = "lm",
              # se=F,
              formula = y~poly(x,2),
              aes(group = Ntiming,
                  col=Ntiming)) +
  facet_wrap(~stand.age)
  labs(caption = "this is the cleanest dataset for r100, and it really shows no difference from adding N or obvious trend from timing")


```


# Anova tables
```{r}
## Yield
lme.r100.all.int<- lmer(yield.kgperha~Napplied_update*Ntiming*stand.age+(1|Block), data=r100)
#summary(lme.r100.all.int)
#anova(lme.r100.all.int)
#ref_grid(lme.r100.all.int)
#confint(lme.r100.all.int)
Anova(lme.r100.all.int, type="II")

## Lodging
lme.r100.lodge<- lmer(Lodging~Napplied_update*Ntiming+(1|Block), data=r100)
#summary(lme.r100.lodge)
#anova(lme.r100.lodge)
#ref_grid(lme.r100.lodge)
#confint(lme.r100.lodge)
Anova(lme.r100.lodge, type="II")

## Height
r100.long<- pivot_longer(data=r100, 
                                cols=Height_1:Height_5, 
                                names_to = "Height", 
                                values_to = "ht.msmt")
str(r100.long)
lme.r100.ht.long<- lmer(ht.msmt~Napplied_update*Ntiming+(1|Block), data=r100.long)
#ref_grid(lme.r100.ht.long)
#summary(lme.r100.ht.long)
#confint(lme.r100.ht.long)
#anova(lme.r100.ht.long)
Anova(lme.r100.ht.long, type="II")


```


# Generate table of average response variables (yield, height, lodging) between stand age and nitrogen application rate
```{r}
# Get average yield and lodging score for Rosemount R-100 (with three sig. figs.)
r100.nitrogenapplied<- r100 %>% group_by(stand.age, Napplied_update) %>% summarise(
  ntot=n(),
  sd.lodging=signif(sd(Lodging),3),
  mean.lodging=signif(mean(Lodging),3),
  sd.yield=signif(sd(yield.kgperha),3),
  mean.yield=signif(mean(yield.kgperha),3),
  SE.yield=sd.yield/(sqrt(ntot))
)

# Join together means and standard deviation values in a single column
r100.nitrogenapplied$yield <- paste(r100.nitrogenapplied$mean.yield, 
                                r100.nitrogenapplied$sd.yield, sep = " ")
r100.nitrogenapplied$Lodging<- paste(r100.nitrogenapplied$mean.lodging, 
                                 r100.nitrogenapplied$sd.lodging, sep = " ")

# Relevel
r100.nitrogenapplied$Napplied_update<- factor(r100.nitrogenapplied$Napplied_update,
                                        levels=c("0","20","40","60","80","100","120","140"))

# Get average height Rosemount R-100 (with three sig. figs.)
r100.nitrogenapplied.long<- r100.long %>% group_by(Napplied_update, stand.age) %>% summarise(
  number_of_samples=n(),
  sd.Height=signif(sd(ht.msmt),3),
  mean.Height=signif(mean(ht.msmt),3)
)

# Relevel
r100.nitrogenapplied.long$Napplied_update<- factor(r100.nitrogenapplied.long$Napplied_update,
                                        levels=c("0","20","40","60","80","100","120","140"))

# Join together means and standard deviation values in a single column
r100.nitrogenapplied.long$Height<- paste(r100.nitrogenapplied.long$mean.Height,
                                     r100.nitrogenapplied.long$sd.Height, sep = " ")

```


# Figure of yield as a function of nitrogen rate by stand age
```{r}
# Estimating marginal means for Rosemount R-100 differences in grain yield amongst nitrogen application rates.

# Experimental model
lme.r100.experimental<- lmer(yield.kgperha~poly(Napplied_update,2)*stand.age+(1|Block), data=r100)
Anova(lme.r100.experimental, type="II")
summary(lme.r100.experimental)

# Prediction model
pred1<- ggpredict(lme.r100.experimental, terms=c("Napplied_update [all]", "stand.age"))
plot(pred1)


##
##


# Add facet labels
label.standages.st<- as_labeller(c(`1`="Stand age 1",`2`="Stand age 2",`3`="Stand age 3"))
# Add color-blind palette
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#999999", "#0072B2", "#D55E00", "#CC79A7")

# Plot figure
r100.yield.rate.age<- ggplot(r100.nitrogenapplied, aes(x=Napplied_update, y=mean.yield, fill=Napplied_update))+
  facet_grid(.~stand.age, labeller = label.standages.st, scales="free")+
  geom_bar(stat="identity", color="black")+
  geom_errorbar(aes(ymax = mean.yield+SE.yield, ymin = mean.yield-SE.yield), width=0.35)+
  labs(x=(expression("Nitrogen application rate" ~ (lb ~ ac^{-1}))),
       y=(expression("Grain yield" ~ (kg ~ ha^{-1}))),
       caption="standard error shown",
       fill="Nitrogen rate"
       )+
  scale_fill_manual(values=cbbPalette)+
  theme_classic()+
  theme(axis.text.y = element_text(size=12),
       axis.title.x.bottom = element_text(size=11),
       axis.title.y = element_text(size=12),
       axis.text.x = element_text(angle=0, hjust=0.5, size=9),
       legend.text = element_text(size=12),
       legend.title = element_text(size=12),
       legend.position = "right",
       plot.title = element_text(hjust = 0.50, face="bold", size=18),
       plot.subtitle = element_text(hjust = 0.50, size=10),
       strip.text.x = element_text(size=12, angle=0),
       strip.text.y = element_text(size=12, angle=-90))
r100.yield.rate.age

```


# Figure of lodging as a function of nitrogen timing and rate
```{r}
# Get average lodging data
rsmtr100.avg.lodge<- r100 %>% group_by(Ntiming_newlabel, Napplied_update, Ntiming) %>% summarise(
  number_of_samples=n(),
  sd.lo=sd(Lodging, na.rm=T),
  mean.lo=mean(Lodging, na.rm=T),
  SE.lo=sd.lo/(sqrt(number_of_samples))
)

# Relevel nitrogen rate for lodging dataset
rsmtr100.avg.lodge$Napplied_update<- factor(rsmtr100.avg.lodge$Napplied_update,
                                        levels=c("0","20","40","60","80","100","120","140"))


# Plotting the interaction of nitrogen timing and application on lodging
plot.r100.lodge<- ggplot(rsmtr100.avg.lodge, aes(x=Napplied_update, y=mean.lo,
                                             fill=Ntiming))+
  geom_bar(stat="identity", position="dodge", color="black")+
  geom_errorbar(aes(ymax = mean.lo+SE.lo, ymin = mean.lo-SE.lo), 
                position = position_dodge2(width=0.20, padding=0.5, preserve = "total"))+
  labs(x=(expression("Nitrogen application rate" ~ (lb ~ ac^{-1}))),
       y="Average lodging (0-10)",
       color="Nitrogen timing",
       shape="Nitrogen timing")+
  scale_y_continuous(breaks=c(0,2,4,6,8,10),limits=c(0,10))+
  scale_fill_hue(c=45, l=80,labels=c("Unfertilized","Fall","Fall split",
                                "Spring","Spring split"))+
  theme_classic()+
  theme(axis.text.y = element_text(size=12),
       axis.title.x.bottom = element_text(size=11),
       axis.title.y = element_text(size=12),
       axis.text.x = element_text(angle=0, hjust=0.5, size=9),
       legend.text = element_text(size=12),
       legend.title = element_text(size=12),
       legend.position = "right",
       plot.title = element_text(hjust = 0.50, face="bold", size=18),
       plot.subtitle = element_text(hjust = 0.50, size=10),
       strip.text.x = element_text(size=12, angle=0),
       strip.text.y = element_text(size=12, angle=-90))
plot.r100.lodge

```


# Figure of height as a function of nitrogen timing and rate
```{r}

# Get average height data
r100.avg.height<- r100.long %>% group_by(Ntiming_newlabel, Napplied_update, Ntiming) %>% summarise(
  number_of_samples=n(),
  sd.ht=sd(ht.msmt, na.rm=T),
  mean.ht=mean(ht.msmt, na.rm=T),
  SE.ht=sd.ht/(sqrt(number_of_samples))
)

# Plot average height
plot.r100.height.avg<- ggplot(r100.avg.height, 
                              aes(x=Napplied_update, y=mean.ht,fill=Ntiming))+
  geom_bar(stat="identity", position = "dodge")+
  geom_errorbar(aes(ymax = mean.ht+SE.ht, ymin = mean.ht-SE.ht), 
                position = position_dodge2(width=0.20, padding=0.5, preserve = "total"))+
  scale_x_continuous(breaks=c(0,60,80,100,120,140),limits=c(-10,150))+
  labs(x=(expression("Nitrogen application rate" ~ (lb ~ ac^{-1}))),
       y="Average plant height (cm)",
       fill="Nitrogen timing |"
       )+
  scale_fill_hue(c=45, l=80, 
                 labels=c("Unfertilized","Fall","Fall split","Spring","Spring split"))+
  coord_cartesian(ylim=c(135,160))+
  theme_classic()+
  theme(axis.text.y = element_text(size=12),
       axis.title.x.bottom = element_text(size=11),
       axis.title.y = element_text(size=12),
       axis.text.x = element_text(angle=0, hjust=0.5, size=9),
       legend.text = element_text(size=12),
       legend.title = element_text(size=12),
       legend.position = "top",
       plot.title = element_text(hjust = 0.50, face="bold", size=18),
       plot.subtitle = element_text(hjust = 0.50, size=10),
       strip.text.x = element_text(size=12, angle=0),
       strip.text.y = element_text(size=12, angle=-90))
plot.r100.height.avg


```


# Figure of yield plot with all data points facetted by stand age point type and color by stand age
```{r}

# Apply facet labels
timing.labels<- as_labeller(c(`control`="Unfertilized",
                              `Fall`="Fall",
                              `Fall split`="Fall split",
                              `Spring`="Spring",
                              `Spring split`="Spring split"))

label.standages.st<- as_labeller(c(`1`="Stand age 1",`2`="Stand age 2",`3`="Stand age 3"))


# Add color-blind palette with black:
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#999999", "#0072B2", "#D55E00", "#CC79A7")

plotr100.yield.all<- ggplot(r100, aes(x=Napplied_update, y=yield.kgperha, color=Ntiming_newlabel, shape=stand.age))+
  facet_grid(.~stand.age, labeller = label.standages.st)+
  geom_point(size=2)+
  labs(x=(expression("Nitrogen application rate" ~ (lb ~ ac^{-1}))),
       y=(expression("Grain yield" ~ (kg ~ ha^{-1}))),
       color="Nitrogen timing",
       shape="Stand age"
  )+
  scale_colour_manual(values=cbbPalette,
                      breaks=c("control","Unfertilized (Fall)","Fall","Fall split","Split (Fall split)","Spring","Spring split"),
                      labels=c("Unfertilized","Unfertilized (Fall)","Fall","Fall split","Split (Fall split)","Spring","Spring split"))+
  scale_x_continuous(breaks=c(0,20,40,60,80,100,120,140))+
  theme_classic()+
  theme(axis.text.y = element_text(size=12),
        axis.title.x.bottom = element_text(size=12),
        axis.title.y = element_text(size=12),
        axis.text.x = element_text(angle=0, hjust=0.5, size=9),
        legend.text = element_text(size=12),
        legend.title = element_text(size=12),
        legend.position = "right",
        plot.title = element_text(hjust = 0.50, face="bold", size=14),
        plot.subtitle = element_text(hjust = 0.50, size=10),
        strip.text.x = element_text(size=12, angle=0),
        strip.text.y = element_text(size=12, angle=-90),
        panel.spacing = unit(1, "lines"))
plotr100.yield.all


```


# Yield with nitrogen timing barplots
```{r}

# Get summary information for yield data at each timing and stand age
r100.yield.timing<- r100 %>% group_by(Ntiming_newlabel, stand.age) %>% summarise(
  ntot=n(),
  sd.yield=sd(yield.kgperha),
  mean.yield=mean(yield.kgperha),
  SE.yield=sd.yield/(sqrt(ntot))
)

# Facet labels
stand.age<- as_labeller(c(`2`="Stand age 2",`3`="Stand age 3"))

# Adjust color-blind palette
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#999999", "#0072B2", "#D55E00")

# Relevel
r100.yield.timing$Ntiming_newlabel<- factor(r100.yield.timing$Ntiming_newlabel,
                                            levels=c("Unfertilized","Unfertilized (Fall)","Fall","Fall split","Split (Fall split)","Spring","Spring split"))

# Plot bar plot
plot.r100.yield.timing<- ggplot(r100.yield.timing, 
                                aes(x=Ntiming_newlabel, y=mean.yield, fill=Ntiming_newlabel))+
  facet_grid(.~stand.age,labeller=stand.age, scales="free")+
  geom_bar(stat="identity", position="dodge")+
  geom_errorbar(aes(ymin = mean.yield-SE.yield, ymax = mean.yield+SE.yield), 
                width = 0.35, color="grey20")+
  labs(x="Nitrogen timing",
       y=(expression("Average grain yield" ~ (kg ~ ha^{-1}))),
       fill="Nitrogen timing")+
  scale_y_continuous(breaks = c(0,200,400,600,800), limits = c(0,700))+
  scale_fill_manual(values=cbbPalette,
                    breaks=c("Unfertilized","Unfertilized (Fall)","Fall","Fall split","Split (Fall split)","Spring","Spring split"),
                    labels=c("Unfertilized","Unfertilized (Fall)","Fall","Fall split","Split (Fall split)","Spring","Spring split"))+
  theme_classic()+
  theme(axis.text.y = element_text(size=12),
        axis.title.x.bottom = element_text(size=12),
        axis.title.y = element_text(size=12),
        axis.text.x = element_text(angle=20, hjust=1, size=10),
        legend.text = element_text(size=12),
        legend.title = element_text(size=12),
        legend.position = "right",
        plot.title = element_text(hjust = 0.50, face="bold", size=14),
        plot.subtitle = element_text(hjust = 0.50, size=10),
        strip.text.x = element_text(size=12, angle=0),
        strip.text.y = element_text(size=12, angle=-90))
plot.r100.yield.timing

```


# Save high-quality figures to working directory
```{r}

# Lodging
#tiff(filename="~/Desktop/Jungers lab/N-rate Timing/Figures/r100.lodging.nrate.timing.tif",res = 600, width = 9, height = 6, units = "in", compression="lzw")
plot.r100.lodge
#dev.off()

# Height
#tiff(filename="~/Desktop/Jungers lab/N-rate Timing/Figures/r100.ht.nrate.timing.tif",res = 600, width = 9, height = 6, units = "in", compression="lzw")
plot.r100.height.avg
#dev.off()

# Grain yield as a function of nitrogen rate and stand age
#tiff(filename="~/Desktop/Jungers lab/N-rate Timing/Figures/r100.yield.nrate.tif",res = 600, width = 9, height = 6, units = "in", compression="lzw")
plotr100.yield.all
#dev.off()

# Grain yield as a function of nitrogen timing and stand age
#tiff(filename="~/Desktop/Jungers lab/N-rate Timing/Figures/r100.yield.timing.tif",res = 600, width = 9, height = 6, units = "in", compression="lzw")
plot.r100.yield.timing
#dev.off()

```





