---
title: "Staples analysis"
author: "Dominic Christensen"
date: "06/17/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---


# Set working directory, load pacakges, load data files, and modify character codes
```{r}

# Load packages
library(readxl)
library(tidyverse)
library(emmeans)
library(lme4)
library(lattice)
library(patchwork)
library(car)
library(patchwork)
library(multcompView)

# Load in staples datasets for analysis
staples.split.dataset<- read.csv("Staples.Nsplit.analysis.only.csv") 
staples.cumulative.dataset<- read.csv("Staples.csv")

# Create dataset for fall & spring only analysis
staples.fallspringonly<- subset(staples.cumulative.dataset, 
                                Ntiming %in% c("Unfertilized","Fall","Spring"))

# Modify character codes
## Staples split
str(staples.split.dataset)
colnames(staples.split.dataset)
names<- c(1,21:23)
staples.split.dataset[,names] <- lapply(staples.split.dataset[,names] , factor)
staples.split.dataset$Lodging<- as.integer(staples.split.dataset$Lodging)

## Staples cumulative
str(staples.cumulative.dataset)
colnames(staples.cumulative.dataset)
names<- c(1,21:24)
staples.cumulative.dataset[,names] <- lapply(staples.cumulative.dataset[,names] , factor)
staples.cumulative.dataset$Lodging<- as.integer(staples.cumulative.dataset$Lodging)

# Staples fall & spring only fertilizer applications
str(staples.fallspringonly)
staples.fallspringonly$Ntiming<- as.factor(staples.fallspringonly$Ntiming)
staples.fallspringonly$Block<- as.factor(staples.fallspringonly$Block)
staples.fallspringonly$stand.age<- as.factor(staples.fallspringonly$stand.age)

```


# Staples split anova tables
```{r}

# Yield
lme.staples.split<- lmer(yield.kgperha~poly(summerNrate,2)*Ntiming*stand.age+
                        (1|Block), data=staples.split.dataset)
Anova(lme.staples.split, type="II")
#summary(lme.staples.split)
#plot(lme.staples.split, main="Yield with predictor variables residuals")


# Lodging
lme.staples.split.lodging<- lmer(Lodging~poly(summerNrate,2)*Ntiming*stand.age+
                        (1|Block), data=staples.split.dataset)
Anova(lme.staples.split.lodging, type="II")
#summary(lme.staples.lodging.all.interactions.poly)
#plot(lme.staples.split.lodging, main="Yield with predictor variables residuals")


# Height
staples.split.dataset.long<- pivot_longer(data=staples.split.dataset, 
                                cols=Height_1:Height_5, 
                                names_to = "Height", 
                                values_to = "ht.msmt")
# Modify character codes
str(staples.split.dataset.long)
staples.split.dataset.long$ht.msmt<- as.numeric(staples.split.dataset.long$ht.msmt)
# Model
lme.staples.split.ht<- lmer(ht.msmt~poly(summerNrate,2)*Ntiming*stand.age+
                        (1|Block), data=staples.split.dataset.long)
Anova(lme.staples.split.ht, type="II")
#summary(staples.ht.long)
#plot(lme.staples.split.ht, main="Yield with predictor variables residuals")


```


# Staples fall & spring only anova tables
```{r}
str(staples.fallspringonly)
str(staples.split.dataset)

# Set as numeric variable
staples.fallspringonly$Lodging<- as.numeric(staples.fallspringonly$Lodging)

# Yield
lme.staples.fs<- lmer(yield.kgperha~poly(updatedNrate,2)*Ntiming*stand.age+
                        (1|Block), data=staples.fallspringonly)
Anova(lme.staples.fs, type="II")
#summary(lme.staples.fs)
#plot(lme.staples.fs, main="Yield with predictor variables residuals")


# Lodging
lme.staples.fs.lodging<- lmer(Lodging~poly(updatedNrate,2)*Ntiming*stand.age+
                        (1|Block), data=staples.fallspringonly)
Anova(lme.staples.fs.lodging, type="II")
#summary(lme.staples.fs.lodging)
#plot(lme.staples.fs.lodging, main="Yield with predictor variables residuals")


# Height
staples.fs.dataset.long<- pivot_longer(data=staples.fallspringonly, 
                                cols=Height_1:Height_5, 
                                names_to = "Height", 
                                values_to = "ht.msmt")
# Modify character codes
str(staples.fs.dataset.long)
staples.fs.dataset.long$ht.msmt<- as.numeric(staples.fs.dataset.long$ht.msmt)

# Model
lme.staples.fs.ht<- lmer(ht.msmt~poly(updatedNrate,2)*Ntiming*stand.age+
                        (1|Block), data=staples.fs.dataset.long)
Anova(lme.staples.fs.ht, type="II")
#summary(lme.staples.fs.ht)
#plot(lme.staples.fs.ht, main="Yield with predictor variables residuals")

```


# Staples timing model (for bar plots)
```{r}
# Model
str(staples.cumulative.dataset)
lme.staples.base.model<- lmer(yield.kgperha~Ntiming+stand.age+
                                (Ntiming:stand.age)+
                                (1|Block), data=staples.cumulative.dataset)
Anova(lme.staples.base.model, type="II")
#summary(lme.staples.base.model)

```


# Staples split dataset modeling grain yield & plotting prediction curves (model & split plot)
```{r}
# Model
lme.staples.split.model<- lmer(yield.kgperha~poly(summerNrate,2)+Ntiming+stand.age+
                          (poly(summerNrate,2):stand.age)+
                          (Ntiming:stand.age)+
                          (1|Block), data=staples.split.dataset)
Anova(lme.staples.split.model, type="II")

# Develop prediction curve
staples.split.model.dataset<- expand.grid(
            summerNrate=c(0,10,20,30,40,50,60,70,80),
            Ntiming=c("control","Fall split","Spring split"),
            stand.age=c(1,2,3),
            Block=c(1,2,3)
)

# Modify character codes
staples.split.model.dataset<- as.data.frame(staples.split.model.dataset)
str(staples.split.model.dataset)
staples.split.model.dataset$Block<- as.factor(staples.split.model.dataset$Block)
staples.split.model.dataset$stand.age<- as.factor(staples.split.model.dataset$stand.age)
staples.split.model.dataset$Ntiming<- as.factor(staples.split.model.dataset$Ntiming)
staples.split.model.dataset$summerNrate<- as.numeric(staples.split.model.dataset$summerNrate)

# Calculate prediction curves
str(staples.split.dataset)
str(staples.split.model.dataset)

str(p3 <- predict(lme.staples.split.model,
                  staples.split.model.dataset,re.form=NA)) # new data, level-0
staples.split.model.dataset<- cbind(staples.split.model.dataset, p3) # no random effects

# Subset for block 1 only (since random effects aren't counted)
staples.split.model.dataset<- subset(staples.split.model.dataset, Block=="1")

# Subset out control line, since it really is not necessary
staples.split.model.dataset<- subset(staples.split.model.dataset, Ntiming!="control")

# Facet grid labels:
label.standages.st<- as_labeller(c(`1`="Stand age 1",`2`="Stand age 2",`3`="Stand age 3"))


# Relabel nitrogen timing column in both dataset such that for stand age 1 fall split is renamed "split (fall split)"
staples.split.dataset$Ntiming_modify<- 
ifelse(staples.split.dataset$Ntiming=="Fall split" & staples.split.dataset$stand.age==1, "Split (Fall split)", 
ifelse(staples.split.dataset$Ntiming=="Spring split", "Spring split",
ifelse(staples.split.dataset$Ntiming=="control","control",
ifelse(staples.split.dataset$Ntiming=="Fall split" & staples.split.dataset$stand.age!=1, "Fall split", "missing"))))

staples.split.model.dataset$Ntiming_modify<- 
ifelse(staples.split.model.dataset$Ntiming=="Fall split" & staples.split.model.dataset$stand.age==1, "Split (Fall split)", 
ifelse(staples.split.model.dataset$Ntiming=="Spring split", "Spring split",
ifelse(staples.split.model.dataset$Ntiming=="control","control",
ifelse(staples.split.model.dataset$Ntiming=="Fall split" & staples.split.model.dataset$stand.age!=1, "Fall split", "missing"))))


# Plot prediction results
plot.staples.split.model.dataset<- ggplot()+
   geom_point(data=staples.split.dataset, 
              aes(x=summerNrate, y=yield.kgperha,color=Ntiming_modify), size=2)+
  geom_line(data=staples.split.model.dataset, 
            aes(x=summerNrate,y=p3, color=Ntiming_modify,linetype=Ntiming_modify),size=1.5)+
  facet_grid(.~stand.age, labeller = label.standages.st)+
  labs(x=(expression("Split nitrogen application rate" ~ (lb ~ ac^{-1}))),
       y=(expression("Grain yield" ~ (kg ~ ha^{-1}))),
       color="Nitrogen timing",
       linetype="Nitrogen timing"
       )+
  scale_color_discrete(
                     breaks=c("Split (Fall split)","Fall split","Spring split"),
                     labels=c("Split (Fall split)","Fall split","Spring split"))+
  scale_linetype_discrete(
                        breaks=c("Split (Fall split)","Fall split","Spring split"),
                        labels=c("Split (Fall split)","Fall split","Spring split"))+
  theme_classic()+
  theme(axis.text.y = element_text(size=8),
       axis.title.x.bottom = element_text(size=12),
       axis.title.y = element_text(size=12),
       axis.text.x = element_text(angle=0, hjust=0.5, size=7),
       legend.text = element_text(size=10),
       legend.title = element_text(size=10),
       legend.position = "right",
       plot.title = element_text(hjust = 0.50, face="bold", size=12),
       plot.subtitle = element_text(hjust = 0.50, size=10),
       strip.text.x = element_text(size=10, angle=0),
       strip.text.y = element_text(size=10, angle=-90),
       panel.spacing = unit(1, "lines"))
plot.staples.split.model.dataset +
  labs(caption = "what are red dots? -JPB")


ggplot()+
   geom_point(data=subset(staples.split.dataset, Ntiming_modify != "control"), 
              aes(x=summerNrate, y=yield.kgperha,color=Ntiming_modify), size=2) +
  geom_line(data=staples.split.model.dataset, 
            aes(x=summerNrate,y=p3, color=Ntiming_modify,linetype=Ntiming_modify),size=1.5)+
  facet_grid(.~stand.age, labeller = label.standages.st)+
  labs(x=(expression("Split nitrogen application rate" ~ (lb ~ ac^{-1}))),
       y=(expression("Grain yield" ~ (kg ~ ha^{-1}))),
       color="Nitrogen timing",
       linetype="Nitrogen timing"
       )+
  scale_color_discrete(
                     breaks=c("Split (Fall split)","Fall split","Spring split"),
                     labels=c("Split (Fall split)","Fall split","Spring split"))+
  scale_linetype_discrete(
                        breaks=c("Split (Fall split)","Fall split","Spring split"),
                        labels=c("Split (Fall split)","Fall split","Spring split"))+
  theme_classic()+
  theme(axis.text.y = element_text(size=8),
       axis.title.x.bottom = element_text(size=12),
       axis.title.y = element_text(size=12),
       axis.text.x = element_text(angle=0, hjust=0.5, size=7),
       legend.text = element_text(size=10),
       legend.title = element_text(size=10),
       legend.position = "right",
       plot.title = element_text(hjust = 0.50, face="bold", size=12),
       plot.subtitle = element_text(hjust = 0.50, size=10),
       strip.text.x = element_text(size=10, angle=0),
       strip.text.y = element_text(size=10, angle=-90),
       panel.spacing = unit(1, "lines")) +
  labs(title = "KEEP - JPB",
       caption = "staples") -> gg1
gg1


```


# Staples prediction plot of grain yield differences from nitrogen rates among stand ages (figure)
```{r}
str(staples.cumulative.dataset)

# Model
lme.staples.overall.model<- lmer(yield.kgperha~poly(updatedNrate,2)*stand.age+
                          (1|Block), data=staples.cumulative.dataset)
Anova(lme.staples.overall.model, type="II")

# Develop prediction curve
staples.overall.model.dataset<- expand.grid(
            updatedNrate=c(0,10,20,30,40,50,60,70,80,90,100,110,120,130,140),
            stand.age=c(1,2,3),
            Block=c(1,2,3)
)

# Modify character codes
staples.overall.model.dataset<- as.data.frame(staples.overall.model.dataset)
str(staples.overall.model.dataset)
staples.overall.model.dataset$Block<- as.factor(staples.overall.model.dataset$Block)
staples.overall.model.dataset$stand.age<- as.factor(staples.overall.model.dataset$stand.age)
staples.overall.model.dataset$updatedNrate<- as.numeric(staples.overall.model.dataset$updatedNrate)

# Calculate prediction curves
str(staples.cumulative.dataset)
str(staples.overall.model.dataset)

str(p3 <- predict(lme.staples.overall.model,
                  staples.overall.model.dataset,re.form=NA)) # new data, level-0
staples.overall.model.dataset<- cbind(staples.overall.model.dataset, p3) # no random effects

# Subset for block 1 only (since random effects aren't counted)
staples.overall.model.dataset<- subset(staples.overall.model.dataset, Block=="1")


# Plot:

# Facet labels
label.standages.st<- as_labeller(c(`1`="Stand age 1",`2`="Stand age 2",`3`="Stand age 3"))

# Plot
plot.staples.overall.model<- ggplot()+
   geom_point(data=staples.cumulative.dataset, aes(x=updatedNrate, y=yield.kgperha,
                                              color=stand.age), size=2)+
  geom_line(data=staples.overall.model.dataset, aes(x=updatedNrate,y=p3,
                                        color=stand.age),size=1.5)+
  facet_grid(.~stand.age, labeller = label.standages.st)+
  labs(x=(expression("Nitrogen application rate" ~ (lb ~ ac^{-1}))),
       y=(expression("Grain yield" ~ (kg ~ ha^{-1}))),
       color="Stand age"
       )+
  scale_x_continuous(breaks = c(0,20,40,60,80,100,120,140), limits = c(-1,145))+
  theme_classic()+
  theme(axis.text.y = element_text(size=7),
       axis.title.x.bottom = element_text(size=10),
       axis.title.y = element_text(size=10),
       axis.text.x = element_text(angle=0, hjust=0.5, size=7),
       legend.text = element_text(size=10),
       legend.title = element_text(size=10),
       legend.position = "right",
       plot.title = element_text(hjust = 0.50, face="bold", size=12),
       plot.subtitle = element_text(hjust = 0.50, size=10),
       strip.text.x = element_text(size=10, angle=0),
       strip.text.y = element_text(size=10, angle=-90),
       panel.spacing = unit(1, "lines")) +
  labs(title = "KEEP - JPB")
plot.staples.overall.model
gg2 <- plot.staples.overall.model


# Get maximum yield values from prediction curve
staples.overall.model.dataset.maximum.values<- staples.overall.model.dataset %>% group_by(stand.age) %>%
  summarise(
    ntot=n(),
    max.yield=max(p3)
  )

```


# Staples summary tables for tables and figures (tables & figures)
```{r}
staples.timing.age<- staples.cumulative.dataset %>% group_by(Ntiming_actual, stand.age) %>%
  summarise(number_of_samples=n(),
            sd.yield=sd(yield.kgperha),
            mean.yield=mean(yield.kgperha),
            SE.yield=sd.yield/(sqrt(number_of_samples))
            )

staples.timing.age2<- staples.cumulative.dataset %>% group_by(Ntiming, stand.age) %>%
  summarise(number_of_samples=n(),
            sd.yield=sd(yield.kgperha),
            mean.yield=mean(yield.kgperha),
            SE.yield=sd.yield/(sqrt(number_of_samples))
            )

staples.nrate.age<- staples.cumulative.dataset %>% group_by(updatedNrate, stand.age) %>%
  summarise(number_of_samples=n(),
            sd.yield=signif(sd(yield.kgperha),3),
            mean.yield=signif(mean(yield.kgperha),3)
            )

# Paste mean and standard deviations together
staples.nrate.age$yield <- paste(staples.nrate.age$mean.yield, staples.nrate.age$sd.yield, sep = " ")


```


# Staples average grain yield bar plot with different nitrogen timings (bar plot)
```{r}
staples.timing.age # data for bar plot
lme.staples.base.model # model for means comparison

# Modify reference grid
staples.regrid.not.control<- ref_grid(lme.staples.base.model, 
                                      at=list(Ntiming=c("Fall","Fall split",
                                                        "Spring","Spring split")))

# Run emmeans
ref_grid(lme.staples.base.model)

staples.timing.emmean<- emmeans(staples.regrid.not.control, ~Ntiming|stand.age)
staples.timing.emmean

# Calculate pairwise contrasts
contrast(staples.timing.emmean, alpha = 0.05, method = "pairwise")
# Get significant letters
a<- multcomp::cld(staples.timing.emmean, alpha=0.05, Letters=LETTERS)

# Linear prediction of yield based on nitrogen timing and stand age
emmip(lme.staples.base.model, Ntiming~stand.age)

# Significant letters
staples.timing.age$sletters<- c( "c","c","bc",
                                 "bc","b","c",
                                 "ab","a","bc",
                                 "a","ab","",
                                 "","","a"
                                 )

# Relevel data set
staples.timing.age$Ntiming_actual<- factor(staples.timing.age$Ntiming_actual,
levels=c("Unfertilized","Unfertilized (Fall)","Fall","Fall split","Split (Fall split)","Spring","Spring split"))

# Plot
plot.staples.timing<- ggplot(staples.timing.age, 
                             aes(x=Ntiming_actual, y=mean.yield, fill=Ntiming_actual))+
  geom_bar(stat="identity", position = "dodge")+
  geom_errorbar(aes(ymin = mean.yield-SE.yield, ymax = mean.yield+SE.yield), 
                width = 0.35, color="grey20")+
  facet_grid(.~stand.age,labeller=label.standages.st, scales="free")+
  geom_text(aes(label = staples.timing.age$sletters, y=mean.yield+SE.yield), vjust = -0.5)+
  labs(x="Nitrogen timing",
       y=(expression("Average grain yield" ~ (kg ~ ha^{-1}))),
       fill="Nitrogen timing")+
  scale_y_continuous(breaks = c(0,200,400,600,800,1000), limits = c(0,1000))+
    scale_fill_hue(c=45, l=80, 
                 labels=c("Unfertilized","Unfertilized (Fall)","Fall","Fall split","Split (Fall split)","Spring","Spring split"))+
  theme_classic()+
  theme(axis.text.y = element_text(size=12),
        axis.title.x.bottom = element_text(size=12),
        axis.title.y = element_text(size=12),
        axis.text.x = element_text(angle=20, hjust=1, size=10),
        legend.text = element_text(size=12),
        legend.title = element_text(size=12),
        legend.position = "right",
        plot.title = element_text(hjust = 0.50, face="bold", size=14),
        plot.subtitle = element_text(hjust = 0.50, size=10),
        strip.text.x = element_text(size=12, angle=0),
        strip.text.y = element_text(size=12, angle=-90)) +
  labs(title = "KEEP -JPB",
       caption = "too many nitrogen timing treatments though, need to adjust")
plot.staples.timing -> gg3
gg3


```


# Staples accumulated grain yield and nitrogen figure (figure)
```{r}

# Plot 1
plot.staples.cumulative.sum<- ggplot(staples.cumulative.dataset,
                         aes(x=updatedNcumsum,y=yield.kgperha,color=Ntiming_actual))+
  facet_grid(.~stand.age, scales = "free")+
  geom_point(size=2)+
  geom_smooth(method="lm", formula = y~poly(x,2),se=F)+ # Applying this function just to quickly see the trendline
  labs(
    x="Nitrogen cumulative sum (lbs/ac)",
    y="Yield (kg/ha)",
    title="Yield responses with stand age, nitrogen timing, and nitrogen rate"
  )+
  theme_bw()+
  theme(axis.text.y = element_text(size=10),
       axis.title.x.bottom = element_text(size=12),
       axis.title.y = element_text(size=12),
       axis.text.x = element_text(angle=0, hjust=0.5, size=11),
       legend.text = element_text(size=10),
       legend.title = element_text(size=12),
       legend.position = "right",
       plot.title = element_text(hjust = 0.50, face="bold", size=12),
       plot.subtitle = element_text(hjust = 0.50, size=10),
       strip.text.x = element_text(size=12, angle=0),
       strip.text.y = element_text(size=12, angle=-90))
plot.staples.cumulative.sum

##
##
##
# Facet grid labels:
label.standages.st<- as_labeller(c(`1`="Stand age 1",`2`="Stand age 2",`3`="Stand age 3"))

# Relevel data set
staples.cumulative.dataset$Ntiming_actual<- factor(staples.cumulative.dataset$Ntiming_actual,
levels=c("Unfertilized","Unfertilized (Fall)","Fall","Fall split","Split (Fall split)","Spring","Spring split"))

# Plot 2
plot.staples.cumulative.sum2<- ggplot(staples.cumulative.dataset,
                         aes(x=updatedNcumsum,y=cumulative.grain.yield,
                             color=Ntiming_actual, shape=Ntiming_actual))+
  facet_grid(.~stand.age, scales = "free", labeller = label.standages.st)+
  geom_point(size=2)+
  #geom_smooth(method="lm", formula = y~poly(x,2), se=F, size=0.5)+ # See trendline quickly
    labs(
       y=(expression("Cumulative grain yield" ~ (kg ~ ha^{-1}))),
       x=(expression("Cumulative nitrogen applied" ~ (lbs ~ ac^{-1}))),
       color="Nitrogen timing",
       shape="Nitrogen timing"
       #title="Cumulative grain yield by stand age and cumulative nitrogen applied"
                 )+
    scale_color_hue( 
                 labels=c("Unfertilized","Unfertilized (Fall)","Fall","Fall split",
                          "Split (Fall split)","Spring","Spring split")
                 )+
    scale_shape_manual(values = c(1,2,3,4,5,6,7), 
                       labels=c("Unfertilized","Unfertilized (Fall)","Fall","Fall split",
                                "Split (Fall split)","Spring","Spring split")
                 )+
  theme_classic()+
  theme(axis.text.y = element_text(size=10),
       axis.title.x.bottom = element_text(size=12),
       axis.title.y = element_text(size=12),
       axis.text.x = element_text(angle=0, hjust=0.5, size=11),
       legend.text = element_text(size=10),
       legend.title = element_text(size=12),
       legend.position = "right",
       plot.title = element_text(hjust = 0.50, face="bold", size=12),
       plot.subtitle = element_text(hjust = 0.50, size=10),
       strip.text.x = element_text(size=12, angle=0),
       strip.text.y = element_text(size=12, angle=-90))
plot.staples.cumulative.sum2



# Plot 3
plot.staples.cumulative.sum3<- ggplot(staples.cumulative.dataset,
                         aes(x=updatedNcumsum,y=cumulative.grain.yield, color=stand.age))+
  facet_grid(.~stand.age, scales = "free")+
  geom_point(size=2)+
  geom_smooth(method="lm", formula = y~poly(x,2), se=F)+ # Applying this function just to quickly see the trendline
  labs(
    x="Cumulative nitrogen applied (lbs/ac)",
    y="Cumulative grain yield (kg/ha)",
    title="Cumulative grain yield increases over stand age and nitrogen rate"
  )+
  theme_bw()+
  theme(axis.text.y = element_text(size=10),
       axis.title.x.bottom = element_text(size=12),
       axis.title.y = element_text(size=12),
       axis.text.x = element_text(angle=0, hjust=0.5, size=11),
       legend.text = element_text(size=10),
       legend.title = element_text(size=12),
       legend.position = "right",
       plot.title = element_text(hjust = 0.50, face="bold", size=12),
       plot.subtitle = element_text(hjust = 0.50, size=10),
       strip.text.x = element_text(size=12, angle=0),
       strip.text.y = element_text(size=12, angle=-90)) +
  labs(title = "KEEP - JPB",
       caption = "I like how relationship gets more quadratic as stand age increases")
plot.staples.cumulative.sum3 -> gg4
gg4


```


Save high-quality figures to working directory
```{r}

#tiff(filename="~/Desktop/Jungers lab/N-rate Timing/Figures/staples.split,timing.tif",res = 600, width = 6, height = 3, units = "in", compression="lzw")
plot.staples.split.model.dataset
#dev.off()

#tiff(filename="~/Desktop/Jungers lab/N-rate Timing/Figures/staples.overall.model.tif",res = 600, width = 6, height = 3, units = "in", compression="lzw")
plot.staples.overall.model
#dev.off()

#tiff(filename="~/Desktop/Jungers lab/N-rate Timing/Figures/staples.timing.barplot.tif",res = 600, width = 9, height = 6, units = "in", compression="lzw")
plot.staples.timing
#dev.off()

#tiff(filename="~/Desktop/Jungers lab/N-rate Timing/Figures/staples.cumulative.sum.tif",res = 600, width = 9, height = 6, units = "in", compression="lzw")
plot.staples.cumulative.sum2
#dev.off()

```








