---
title: overflow from n-rate timing report
author: jesse puka-beals
---

```{r}

```

## treatment table

| location | harvest | timing[^1] | N applied prior to harvest |
|----------|---------|------------|----------------------------|
| V17      | 1       | fall       | no                         |
| V17      | 1       | spring     | yes                        |
| V17      | 1       | summer     | no                         |
| V17      | 2       | fall       | yes                        |
| V17      | 2       | spring     | yes                        |
| V17      | 2       | summer     | no                         |
| V17      | 3       | fall       | yes                        |
| V17      | 3       | spring     | yes                        |
| V17      | 3       | summer     | no                         |
| Staples  | 1       | fall       | no                         |
| Staples  | 1       | spring     | yes                        |
| Staples  | 1       | summer     | yes                        |
| Staples  | 2       | fall       | yes                        |
| Staples  | 2       | spring     | yes                        |
| Staples  | 2       | summer     | yes                        |
| Staples  | 3       | fall       | yes                        |
| Staples  | 3       | spring     | yes                        |
| Staples  | 3       | summer     | yes                        |
| R100     | 1       | fall       | no                         |
| R100     | 1       | spring     | yes                        |
| R100     | 1       | summer     | yes                        |
| R100     | 2       | fall       | yes                        |
| R100     | 2       | spring     | yes                        |
| R100     | 2       | summer     | yes                        |
| R100     | 3       | fall       | yes                        |
| R100     | 3       | spring     | yes                        |
| R100     | 3       | summer     | yes                        |

[^1]: fall timing is applied in year prior to harvest

At V17, plots with a spring timing got fertilized prior to the first kernza harvest. If any of that N remains in the soil and rolls over into the next year, it will make spring N plots look higher yielding then they are.

At R100 and Staples, this effect will be diluted by the summer applied N



+-------------------+-------------+--------------+------------+----------------------+-----------------------------------------------+
| balanced datasets | name        | locations    | stand.ages | treatments           | model                                         |
+===================+=============+==============+============+======================+===============================================+
| timing            | dat_timing  | Staples, V17 | 1,2,3      | control, spring only | y\~timing X location X stand.age              |
+-------------------+-------------+--------------+------------+----------------------+-----------------------------------------------+
| timing2           | dat_timing2 | Staples, V17 | 2,3        | control, fall only   | y\~timing X location X stand.age              |
+-------------------+-------------+--------------+------------+----------------------+-----------------------------------------------+
| timing3           |             | R100         | 2,3        | control, spring only |                                               |
+-------------------+-------------+--------------+------------+----------------------+-----------------------------------------------+
| nrate             | dat_nrate   | Staples      | 1,2,3      | all                  | y\~updatednrate X location X stand.age        |
+-------------------+-------------+--------------+------------+----------------------+-----------------------------------------------+
| nrate older stand | dat_old     | R100         | 2,3        | all                  | y\~cum_n\_harvestreset X location X stand.age |
+-------------------+-------------+--------------+------------+----------------------+-----------------------------------------------+
| cumulative yield  |             | Staples, V17 | 1,2,3      | control,spring only  |                                               |
+-------------------+-------------+--------------+------------+----------------------+-----------------------------------------------+
|                   |             |              |            |                      |                                               |
+-------------------+-------------+--------------+------------+----------------------+-----------------------------------------------+

+-------------------+----------------------------------------------------------------------------------------------+
| balanced datasets | testable hypotheses                                                                          |
+===================+==============================================================================================+
| dat_timing        | Yields are similar regardless of fertilization                                               |
|                   |                                                                                              |
|                   | Yields are similar regardless of timing (treatment != control)                               |
|                   |                                                                                              |
|                   | Yields are similar regardless of stand.age or location                                       |
+-------------------+----------------------------------------------------------------------------------------------+
| dat_nrate         | yields are similar regardless of fertilization                                               |
|                   |                                                                                              |
|                   | yields are similar regardless of timing\*rate (treatment != control)                         |
|                   |                                                                                              |
|                   | Yields are similar regardless of stand.age or location                                       |
|                   |                                                                                              |
|                   | Yields are similar regardless of the amount of N applied prior to harvest                    |
|                   |                                                                                              |
|                   | Yield response to N is best modeled as no effect vs a linear, quadratic and quadratic planar |
+-------------------+----------------------------------------------------------------------------------------------+
| dat_old           | Yields in older stands are similar regardless of fertilization                               |
|                   |                                                                                              |
|                   | yields in older stands are similar regardless of timing (treatment != control)               |
|                   |                                                                                              |
|                   | yields in older stands are similar regardless of timing\*rate (treatment != control)         |
|                   |                                                                                              |
|                   | yields in older stands are similar regardless of the amount of N applied prior to harvest    |
|                   |                                                                                              |
|                   | yields are similar regardless of stand age and location                                      |
+-------------------+----------------------------------------------------------------------------------------------+

**Additional questions/framings**

What nitrogen rate results in highest grain yield?

As stand age increases, does optimal nitrogen rate increase?

**Dominic's findings**

R100: timing does not matter, 0-60N maximizes yield

Staples: applying N in fall near 80 lb A is best

V17: fall application is best (including split)

**Jesse's findings**

R100: does not make a lot of sense

# Old analysis

```{r}

dat_v17_tiny %>%
  bind_rows(dat_staples_tiny) -> dat_timing

dat_staples -> dat_nrate

dat_r100 %>%
  bind_rows(dat_staples) %>%
  filter(stand.age != "1") -> dat_old

```

## Dat_timing

```{r no fall}

dat_timing %>%
  filter(timing != "fall") %>%
  lm(yield.kgperha~timing*stand.age*location,
     .
     ) %>%
  # anova()
  emmeans(~timing)

```

We reject Ho that yields are similar between plots that were and were no fertilized in the spring across all 3 years at V17 and R100.

```{r fall and spring but only years 2 and 3}

# we need to add in R100 now

dat_timing %>%
  filter(stand.age != "1") %>%
  lm(yield.kgperha~timing*stand.age*location,
     .
     ) %>%
  anova()
  # emmeans(~timing)


```

```{r}
## Mixed effect models don't work here
# Block explains nothing and locations are too few to treat as random

# lmer(
#   yield.kgperha~timing*stand.age+
#     (1|location/block),
#   dat_timing
# ) %>%
#   car::Anova()


dat_timing %>%
  # distinct(location)
  # filter(location=="Staples") %>%
  filter(location=="RSMT V17") %>%
  lm(
    # yield.kgperha~timing*location*stand.age,
    yield.kgperha~timing*stand.age,
    .
  ) %>%
  # summary()
  anova()
  # emmeans(~timing*stand.age) %>%
  # cld(Letters=letters, reverse=T)

# 3-way interaction with location and stand.age
# staples interaction between timing and stand age

# interaction
dat_timing %>%
  filter(location=="Staples") %>%
  # filter(location=="RSMT V17") %>%
  ggplot(aes(timing,yield.kgperha,
             group = stand.age,
             col=stand.age)) +
  stat_summary(geom = "line")

# This interaction is due to year1 fall essentially being a control

dat_timing %>%
  filter(stand.age != "1" | timing != "fall") %>%
  filter(location == "Staples") %>%
  # ggplot(aes(timing,yield.kgperha,
  #            group = stand.age,
  #            color=stand.age)) +
  # stat_summary(geom = "line")
  lm(
    yield.kgperha~timing*stand.age,
    .
  ) %>%
  anova()


```

| dataset    | location | fixed effects             | p-value |
|------------|----------|---------------------------|---------|
| dat_timing | combined | timing:location:stand.age | \*\*    |
| dat_timing | staples  | timing                    | \*\*    |
| dat_timing | staples  | stand.age                 | \*\*\*  |
| dat_timing | staples  | timing\*stand.age         | \*      |
| dat_timing | v17      | timing                    | \*\*\*  |
| dat_timing | v17      | stand.age                 | \*\*\*  |
| dat_timing | v17      | timing\*stand.age         | NS      |

The effects of timing and stand age were not consistent across sites, and required each site be analyzed separately.

*Yields are similar regardless of fertilization*

Staples: there is an interaction between the effect of fertilization and stand.age.

R100:

Yields are similar regardless of timing (treatment != control)

Yields are similar regardless of stand.age or location

